### 2020-12-30 osf/comps/dataset/raw/README.Rmd

---
title: "Raw data"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

This directory contains `.wav` files of the participants’ production from the experiment.

```{r retrieve}
# Main repo "Many Speech Analysis": https://osf.io/3bmcp/
osf_repo <- osf_retrieve_node("3bmcp")
# Repo components
osf_comps <- osf_ls_nodes(osf_repo)
# Dataset & Materials component
data_comp <- osf_comps %>% filter(name == "Dataset & Materials")
```

```{r data_files, include=TRUE}
osf_ls_files(data_comp, "Raw data") %>%
  select(name) %>%
  filter(str_detect(name, ".*\\.md", negate = TRUE)) %>%
  arrange(name)
```