### 2021-01-05 osf/comps/data/production/trial-lists/README.Rmd

---
title: "Trial lists of the production study"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

This directory contains one `.csv` file per participant from the production experiment with information about the trials (one row per trial).

The files include the following columns:

- `condition`: The condition of the trial (`NF` = noun focus, `AF` = adjective focus, `ANF` = adjective/noun focus).
- `target_name`: Name of the target object.
- `target_colour`: Color of the target object.
- `competitor_name`: Name of the competitor object.
- `competitor_colour`: Color of the competitor object.
- `distractor1_name`: Name of the distractor 1.
- `distractor1_colour`: Color of the distractor 1.
- `distractor2_name`: Name of the distractor 2.
- `distractor2_colour`: Color of the distractor 2.

```{r retrieve}
# Main repo "Many Speech Analysis": https://osf.io/3bmcp/
osf_repo <- osf_retrieve_node("3bmcp")
# Repo components
osf_comps <- osf_ls_nodes(osf_repo)
# Dataset & Materials component
data_comp <- osf_comps %>% filter(name == "Data")
```

```{r data_files, include=FALSE}
osf_ls_files(data_comp, "production/trial-lists") %>%
  select(name) %>%
  filter(str_detect(name, ".*\\.md", negate = TRUE)) %>%
  arrange(name)
```