### 2021-01-31 osf/data/README.Rmd
:bento: - upload TextGrids and READMEs to osf

---
title: "Data component"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

This component hosts the data of the *Many Speech Analyses* project.
For an overview, see the [wiki](https://osf.io/rdtx5/wiki/home/).

There are 2 subdirectories:

- The `norming/` directory holds the files of the norming study.
- The `production/` directory contains two subdirectories:
  - `audio/`: with `.wav` and `.TextGrid` files from the production experiment.
  - `trial-lists/`: with `.csv` files with info on the trials from the production experiment.

See the `README.md` files of each individual directory for more information about the contents.

```{r retrieve}
# Main repo "Many Speech Analysis": https://osf.io/3bmcp/
osf_repo <- osf_retrieve_node("3bmcp")
# Repo components
osf_comps <- osf_ls_nodes(osf_repo)
# Data component
data_comp <- osf_comps %>% filter(name == "Data")
```

```{r data_files, include=TRUE}
osf_ls_files(data_comp) %>%
  select(name) %>%
  filter(str_detect(name, ".*\\.md", negate = TRUE)) %>%
  arrange(name)
```