### 2021-01-31 osf/data/norming/ratings/README.Rmd

---
title: "Typicality ratings of the norming study"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

Here you can find an aggregated dataset with the typicality judgement for each of the images in `norming/images/`, for each participant.

The `tipicality.csv` file has the following columns:

- `prolific id`: The participant/trial id.
- `object`: The name of the object represented in the image.
- `typicality`: The typicality rating (numeric, `1-100`, `1` = most atypical, `100` = most typical).
- `colour`: The color of the object in the image.