### 2021-02-05 scripts/simulations/simulations.Rmd
:construction: - start rmd of RR simulations
You can have a look at the html output to get an idea of what it looks like.

---
title: "Perspective simulation"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    toc_depth: 3
    number_sections: yes
    use_bookdown: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(rmdformats)
library(tidyverse)
library(patchwork)
library(brms)
library(extraDistr)
my_seed <- 115116106
```

# Meta-analysis

## Simulate data set

We first simulate a dataset (`eta_tbl`) of standardised effect sizes.
Each team ($n = 12$) contributes a standardized effect size $\eta_i$, with respective standard error $\text{se}_i$.
The $\eta_i$ is randomly sampled from a $Normal(1, 0.5)$ distribution, while $\text{se}_i$ from a $HalfCauchy(0, 0.1)$.

```{r eta-tbl}
set.seed(my_seed)

# Number of teams, each team contributes one effect size
n <- 12
# standardized effect size
eta <- rnorm(n, 1, 0.5)
# standard error of the standardised effect size
se <- rhcauchy(n, 0.1)
# label each team
team <- letters[1:n]

# put together in a tibble
eta_tbl <- tibble(
  eta,
  se,
  team
)
```

## Model priors

In the meta-analytical model, we will use the following priors:

<!-- TODO: need to discuss prior for sd -->

- A $Normal(0, 1)$ distribution for the prior of the intercept.
- A $HalfCauchy(0, 0.01)$ distribution for the prior of the intercept standard deviation.

```{r meta-priors}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 0.01), class = sd)
)
```

### Prior predictive checks

```{r prior-predictive}
meta_bm_ppred <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  sample_prior = "only",
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm_ppred"
)
```

```{r ppred-samples}
ppred_samples <- posterior_samples(meta_bm_ppred)
glimpse(ppred_samples)
```

## Meta-analytical model

```{r meta-bm}
meta_bm <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm"
)
```

```{r meta-bm-summary}
meta_bm
```

```{r plot-meta-bm}
plot(meta_bm)
```