### 2021-02-09 scripts/simulations/simulations.Rmd
:sparkles: - add analytical and demographic features

---
title: "Perspective simulation"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    toc_depth: 3
    number_sections: yes
    use_bookdown: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(rmdformats)
library(tidyverse)
library(patchwork)
library(brms)
library(extraDistr)
library(betapart)
library(summarytools)
my_seed <- 115116106
```

# Meta-analysis

## Simulate data

We first simulate a dataset (`eta_tbl`) of standardised effect sizes.
Each team ($n = 12$) contributes a standardized effect size $\eta_i$, with respective standard error $\text{se}_i$.

### Helper functions

```{r helpers}
at_least <- function(x) {
  if (1 %in% x) {x} else {sample(c(1, 0, 0, 0, 0))}
}

sdi_mean <- function(preds) {
  bpair <- tibble::tibble(preds) %>% tidyr::unnest_wider(preds, names_sep = ".") %>%
    as.matrix() %>%
    betapart::beta.pair(.)
  sdim <- as.matrix(bpair$beta.sor)
  diag(sdim) <- NA
  rowSums(sdim, na.rm = T) / length(preds)
}
```

### Tibble

The $\eta_i$ is randomly sampled from a $Normal(1, 0.5)$ distribution, while $\text{se}_i$ from a $HalfCauchy(0, 0.1)$.

```{r sim-data}
set.seed(my_seed)

# Number of teams, each team contributes one effect size
n <- 12
# standardized effect size
eta <- rnorm(n, 1, 0.5)
# standard error of the standardised effect size
se <- rhcauchy(n, 0.1)
# label each team
team <- letters[1:n]

# pop-level predictors
pop_pred <- lapply(replicate(n, list(rbinom(5, 1, 0.5))), at_least)
# group-level predictors
group_pred <- lapply(replicate(n, list(rbinom(5, 1, 0.5))), at_least)
# number of post-hoc changes to the measures
phoc_n <- rpois(n, 1)
# number of models run
mod_n <- sample(1:5, n, replace = T)
# major dimension
mdim <- sample(c("duration", "amplitude", "f0", "spectral", "other"), n, TRUE)
# temporal window
twin <- sample(c("segment", "syllable", "word", "phrase"), n, T)
# was data excluded?
excl <- rbinom(n, 1, 0.5)

# research experience as years from PhD
exp <- replicate(n, sample(-3:20, sample(1:10, 1), replace = T))
# prior belief in the research hypothesis
belf <- lapply(lengths(exp), function(.x) rbinom(.x, 1, 0.5))
```

```{r eta-tbl}
# put together in a tibble
eta_tbl <- tibble(
  eta,
  se,
  team,
  pop_pred,
  pop_n = map(pop_pred, sum),
  group_pred,
  group_n = map(group_pred, sum),
  pop_sdi = sdi_mean(pop_pred),
  group_sdi = sdi_mean(group_pred),
  phoc_n,
  mod_n,
  mdim,
  twin,
  excl,
  exp,
  belf
)
```

## Model priors

In the meta-analytical model, we will use the following priors:

<!-- TODO: need to discuss prior for sd -->

- A $Normal(0, 1)$ distribution for the prior of the intercept.
- A $HalfCauchy(0, 0.01)$ distribution for the prior of the intercept standard deviation.

```{r meta-priors}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 0.01), class = sd)
)
```

### Prior predictive checks

```{r prior-predictive}
meta_bm_ppred <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  sample_prior = "only",
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm_ppred"
)
```

```{r ppred-samples}
ppred_samples <- posterior_samples(meta_bm_ppred)
glimpse(ppred_samples)
```

## Meta-analytical model

```{r meta-bm}
meta_bm <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  chain = 1,
  seed = my_seed,
  file = "./models/sim/meta_bm"
)
```

```{r meta-bm-summary}
meta_bm
```

```{r plot-meta-bm}
plot(meta_bm)
```

# Simulated data summary

```{r dfsummary}
# seems dfSummary() does not support list columns
dfSummary(eta_tbl %>% select(-where(is.list)))
```