### 2021-02-22 scripts/r/simulations/simulations.Rmd
:memo: - start factors model

---
title: "Perspective simulation"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    toc_depth: 3
    number_sections: yes
    use_bookdown: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(rmdformats)
library(tidyverse)
theme_set(theme_light())
library(patchwork)
library(brms)
library(extraDistr)
library(HDInterval)
library(betapart)
# library(summarytools)
my_seed <- 115116106
```

# Simulate data

We first simulate a dataset (`eta_tbl`) of standardised effect sizes.
Each team ($n = 12$) contributes a standardized effect size $\eta_i$, with respective standard error $\text{se}_i$.

### Helper functions

```{r helpers}
at_least <- function(x) {
  if (1 %in% x) {x} else {sample(c(1, 0, 0, 0, 0))}
}

sdi_mean <- function(preds) {
  bpair <- tibble::tibble(preds) %>% tidyr::unnest_wider(preds, names_sep = ".") %>%
    as.matrix() %>%
    betapart::beta.pair(.)
  sdim <- as.matrix(bpair$beta.sor)
  diag(sdim) <- NA
  rowSums(sdim, na.rm = T) / length(preds)
}
```

### Tibble

The $\eta_i$ is randomly sampled from a $Normal(1, 0.5)$ distribution, while $\text{se}_i$ from a $HalfCauchy(0, 0.15)$.

```{r sim-data}
set.seed(my_seed)

# Number of teams, each team contributes one effect size
n <- 12
# standardized effect size
eta <- rnorm(n, 1, 0.5)
# standard error of the standardised effect size
se <- rhcauchy(n, 0.15)
# label each team
team <- letters[1:n]

# pop-level predictors
pop_pred <- lapply(replicate(n, list(rbinom(5, 1, 0.5))), at_least)
# group-level predictors
group_pred <- lapply(replicate(n, list(rbinom(5, 1, 0.5))), at_least)
# number of post-hoc changes to the measures
phoc_n <- rpois(n, 1)
# number of models run
mod_n <- sample(1:5, n, replace = T)
# major dimension
mdim <- sample(c("duration", "amplitude", "f0", "spectral", "other"), n, TRUE)
# temporal window
twin <- sample(c("segment", "syllable", "word", "phrase"), n, T)
# was data excluded?
excl <- rbinom(n, 1, 0.5)
# reviewers' ratings
rating <- replicate(n, sample(0:100, sample(3:4, 1), replace = T))

# research experience as years from PhD
exp <- replicate(n, sample(-3:20, sample(1:10, 1), replace = T))
# prior belief in the research hypothesis
belf <- lapply(lengths(exp), function(.x) rbinom(.x, 1, 0.5))
```

```{r eta-tbl}
# put together in a tibble
eta_tbl <- tibble(
  eta,
  se,
  team,
  pop_pred,
  pop_n = map(pop_pred, sum),
  group_pred,
  group_n = map(group_pred, sum),
  pop_sdi = sdi_mean(pop_pred),
  group_sdi = sdi_mean(group_pred),
  phoc_n,
  mod_n,
  mdim,
  twin,
  excl,
  rating,
  exp,
  belf,
  rating_mean = unlist(map(rating, mean)),
  rating_sd = unlist(map(rating, sd)),
  exp_mean = unlist(map(exp, mean)),
  exp_sd = unlist(map(exp, sd)),
  belf_mean = unlist(map(belf, mean)),
  belf_sd = unlist(map(belf, sd))
) %>%
  mutate(
  # need the following in cases where score = 0
  rating_mean = ifelse(rating_mean == 0, 0.01, rating_mean),
  exp_mean = ifelse(exp_mean == 0, 0.01, exp_mean),
  belf_mean = ifelse(belf_mean == 0, 0.01, belf_mean),
  # need the following in cases where there only one person in the team
  rating_sd = ifelse(is.na(rating_sd) | rating_sd == 0, 0.01, rating_sd),
  exp_sd = ifelse(is.na(exp_sd) | exp_sd == 0, 0.01, exp_sd),
  belf_sd = ifelse(is.na(belf_sd) | belf_sd == 0, 0.01, belf_sd)
)
```

# Meta-analytical model

## Model formula

$$
\begin{aligned}
\eta_i      & \sim \text{Normal}(\mu_i, \sigma_i) \\
\mu_i       & = \alpha + \alpha_{\text{t}[i]} \\
\alpha      & \sim \text{Normal}(0, 1) \\
\sigma_{\alpha_{\text{t}}} & \sim \text{HalfCauchy}(0, 1) \\
\sigma_i    & = \text{se}_i
\end{aligned}
$$

## Model priors

In the meta-analytical model, we will use the following priors:

- A $Normal(0, 1)$ distribution for the prior of the intercept.
- A $HalfCauchy(0, 1)$ distribution for the prior of the teams' intercept standard deviation.

```{r sd-prior}
tibble(
  x = seq(0, 10, by = 0.05),
  density = dhcauchy(x, 1)
) %>%
  ggplot(aes(x, density)) +
  geom_area(fill = "#325756", alpha = 0.9) +
  labs(
    title = "HalfCauchy(0, 1)",
    x = "teams' intercept standard deviation"
  )
```

```{r meta-priors}
priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(cauchy(0, 1), class = sd)
)
```

### Prior predictive checks

```{r prior-predictive}
meta_bm_ppred <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  sample_prior = "only",
  chain = 4,
  seed = my_seed,
  file = "./models/sim/meta_bm_ppred"
)

ppred_samples <- posterior_samples(meta_bm_ppred) %>%
  rename(intercept = b_Intercept, `group-level SD` = sd_team__Intercept)
```

```{r ppred-posts}
ppred_samples %>%
  pivot_longer(intercept:`group-level SD`, names_to = "term") %>%
  mutate(term = factor(term, levels = c("intercept", "group-level SD"))) %>%
  ggplot(aes(value)) +
  geom_density(colour = NA, fill = "#325756", alpha = 0.9, bw = 0.4) +
  facet_wrap(~ term, scales = "free")
```

```{r ppred-joint-post}
ppred_samples %>%
  select(intercept, `r_team[a,Intercept]`:`r_team[l,Intercept]`) %>%
  pivot_longer(-intercept, names_to = "team") %>%
  mutate(eta = intercept + value) %>%
  ggplot(aes(eta)) +
  geom_density(colour = NA, fill = "#325756", alpha = 0.9, bw = 0.4) +
  xlim(-7, 7)
```

## Meta-analytical model

```{r meta-bm}
meta_bm <- brm(
  eta | se(se) ~
    (1 | team),
  data = eta_tbl,
  prior = priors,
  chain = 4,
  seed = my_seed,
  cores = 4,
  file = "./models/sim/meta_bm"
)
```

```{r meta-bm-summary}
meta_bm
```

```{r plot-meta-bm}
plot(meta_bm)
```

# Factors analysis model

```{r factors-priors}
factors_priors <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 1), class = b),
  prior(normal(0, 1), class = meanme),
  prior(cauchy(0, 1), class = sdme),
  prior(lkj(2), class = corme)
)
```

```{r factors-bm}
factors_bm <- brm(
  eta | se(se) ~
    pop_sdi +
    phoc_n +
    mod_n +
    mdim +
    twin +
    excl +
    me(rating_mean, rating_sd) +
    me(exp_mean, exp_sd) +
    me(belf_mean, belf_sd),
  data = eta_tbl,
  prior = factors_priors,
  chain = 4,
  iter = 4000,
  control = list(max_treedepth = 15),
  seed = my_seed,
  cores = 4,
  file = "./models/sim/factors_bm"
)
```