### 2021-03-24 osf/data/README.Rmd
:bento: - update osf data; edit RR

---
title: "Data component"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

This component hosts the data of the *Many Speech Analyses* project.
For an overview, see the [wiki](https://osf.io/rdtx5/wiki/home/).

There are 2 subdirectories:

- The `norming/` directory holds the files of the norming study.
  - `images/`: re-colored images for the norming study.
  - `ratings/`: rating results of the norming study.
- The `production/` directory contains two subdirectories:
  - `audio/`: with `.wav` and `.TextGrid` files from the production experiment.
  - `trial-lists/`: with `.csv` files with info on the trials from the production experiment.

See the `README.md` files in each directory for more information about the contents.