### 2021-03-24 osf/data/production/README.Rmd

---
title: "Production study"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

This directory contains the files of the production study, which was designed to elicit spoken utterances that contain referential phrases like *blue banana*.

# Overview

There are 2 subdirectories:

- `audio/` contains `.wav` and `.TextGrid` files from the production experiment.
- `trial-lists/` contains `.csv` files with info on the trials from the production experiment.

# TextGrids

# Trial lists

Trial lists contain the following columns:

- `speaker`: The speaker id.
- `trial`: The position of the trial within the experiment (numeric).
- `condition`: The focus condition (`NF` = noun focus, `AF` = adjective focus, `ANF` = double focus on adjective and noun; `NF`s are the critical conditions).
- `typicality`: Typicality category (`typical`, `medium`, `atypical`; `NA` for AF and ANF).
- `target_object`: Target object.
- `target_colour`: Colour of the target object.
- `competitor_name`: Competitor object.
- `competitor_colour`: Colour of the competitor object.
- `distractror1_name`: Distractor-1 object.
- `distractror1_colour`: Colour of the distractor-1 object.
- `distractror2_name`: Distractor-2 object.
- `distractror2_colour`: Colour of the distractor-2 object.
- `object_en`: The English translation of `target_object`.
- `colour_en`: The English translation of `target_colour`.
- `typ_sd`: The standard deviation of the typicality ratings .
- `typ_median`: The median typicality rating.
- `typ_median`: The median typicality rating.