### 2021-04-15 scripts/r/sample-size/sample-size_TR.Rmd
attempt to alter script

---
title: "Sample size"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
theme_set(theme_light())
library(brms)
library(extraDistr)
library(HDInterval)
library(tidybayes)
```

```{r}
dir.create("./figs/sample-size", FALSE)
```

```{r Silberzahn}

# load in Silberzahn et al. data: https://osf.io/fa743/
silber = read_csv("https://osf.io/fa743/download")

## run model on OR to estimate between teams SD

# set weakly informative priors
# mean OR
round(mean(silber$OR),2)

priors <- c(
  prior(normal(1.39, 1), class = Intercept),
  prior(cauchy(0, 0.5), class = sd)
)

# run simple model
silber_sd <- brm(
  OR ~ (1 | Team),
  data = silber,
  prior = priors,
  chain = 4,
  iter = 4000,
  cores = 4,
  backend = "cmdstanr", threads = threading(4),
  control = list(adapt_delta = 0.99999, max_treedepth = 15),
)

# extract SD estimate and 95 CrI
SD = summary(silber_sd)$random[[1]][1]
lower = summary(silber_sd)$random[[1]][3]
upper = summary(silber_sd)$random[[1]][4]

quantiles <- quantile(seq(lower, upper, by = 0.1), probs = seq(0, 1, 0.25),  na.rm = FALSE)
quantiles <- c(quantiles[[1]], quantiles[[2]], quantiles[[3]], quantiles[[4]], quantiles[[5]])

```

```{r functions}
simulate_data <- function(sd, n) {
  tibble(
    es_mean = rnorm(n, 0, sd),
    # no measurement error model for the simulation
    # es_se = rhcauchy(n, 0.15),
    team = letters[1:n]
  )
}

run_brm <- function(data) {
  priors <- c(
    prior(normal(0, 1), class = Intercept),
    prior(cauchy(0, 0.15), class = sd)
  )

  meta_bm <- brm(
    es_mean ~
    #es_mean | se(es_se) ~
      (1 | team),
    data = data,
    prior = priors,
    chain = 4,
    cores = 4,
    backend = "cmdstanr", threads = threading(4),
    control = list(adapt_delta = 0.9999, max_treedepth = 15),
  )

  return(meta_bm)
}

run_simulations <- function(sd, n = 10) {
  data <- simulate_data(sd, n)
  brm_model <- run_brm(data)
  return(brm_model)
}
```

```{r run-sims}
if (file.exists("./models/sim/sims_40.rds")) {
  sims_20 <- readRDS("./models/sim/sims_20.rds")
  sims_25 <- readRDS("./models/sim/sims_25.rds")
  sims_30 <- readRDS("./models/sim/sims_30.rds")
  sims_35 <- readRDS("./models/sim/sims_35.rds")
  sims_40 <- readRDS("./models/sim/sims_40.rds")
} else {

  sims_20 <- tibble(
    team_sd = quantiles,
    model = map(
      team_sd,
      run_simulations
    )
  )

  saveRDS(sims_20, "./models/sim/sims_20.rds")

  sims_25 <- tibble(
    team_sd = quantiles,
    model = map(
      team_sd,
      ~run_simulations(.x, 25)
    )
  )

  saveRDS(sims_25, "./models/sim/sims_25.rds")

  sims_30 <- tibble(
    team_sd = quantiles,
    model = map(
      team_sd,
      run_simulations(.x, 30)
    )
  )

  saveRDS(sims_30, "./models/sim/sims_30.rds")

  sims_35 <- tibble(
    team_sd = quantiles,
    model = map(
      team_sd,
      ~run_simulations(.x, 35)
    )
  )

  saveRDS(sims_35, "./models/sim/sims_35.rds")

  sims_40 <- tibble(
    team_sd = quantiles,
    model = map(
      team_sd,
      ~run_simulations(.x, 40)
    )
  )

  saveRDS(sims_40, "./models/sim/sims_40.rds")
}
```

```{r team-sds}
sds_20 <- map(
  sims_20$model,
  ~spread_draws(.x, sd_team__Intercept) %>% summarise(sd = sd(sd_team__Intercept)) %>% pull
) %>%
  unlist()

sds_25 <- map(
  sims_25$model,
  ~spread_draws(.x, sd_team__Intercept) %>% summarise(sd = sd(sd_team__Intercept)) %>% pull
) %>%
  unlist()

sds_30 <- map(
  sims_30$model,
  ~spread_draws(.x, sd_team__Intercept) %>% summarise(sd = sd(sd_team__Intercept)) %>% pull
) %>%
  unlist()

sds_35 <- map(
  sims_35$model,
  ~spread_draws(.x, sd_team__Intercept) %>% summarise(sd = sd(sd_team__Intercept)) %>% pull
) %>%
  unlist()

sds_40 <- map(
  sims_40$model,
  ~spread_draws(.x, sd_team__Intercept) %>% summarise(sd = sd(sd_team__Intercept)) %>% pull
) %>%
  unlist()
```

```{r plot}
sd_20 <- tibble(
  team_sd = quantiles,
  sd = sds_20,
  min = sapply(sds_20, function(.x) inverseCDF(0.025, phcauchy, .x)),
  max = sapply(sds_20, function(.x) inverseCDF(0.975, phcauchy, .x)),
  width = max - min,
  n = 20
)

sd_25 <- tibble(
  team_sd = quantiles,
  sd = sds_25,
  min = sapply(sds_25, function(.x) inverseCDF(0.025, phcauchy, .x)),
  max = sapply(sds_25, function(.x) inverseCDF(0.975, phcauchy, .x)),
  width = max - min,
  n = 25
)

sd_30 <- tibble(
  team_sd = quantiles,
  sd = sds_30,
  min = sapply(sds_30, function(.x) inverseCDF(0.025, phcauchy, .x)),
  max = sapply(sds_30, function(.x) inverseCDF(0.975, phcauchy, .x)),
  width = max - min,
  n = 30
)

sd_35 <- tibble(
  team_sd = quantiles,
  sd = sds_35,
  min = sapply(sds_35, function(.x) inverseCDF(0.025, phcauchy, .x)),
  max = sapply(sds_35, function(.x) inverseCDF(0.975, phcauchy, .x)),
  width = max - min,
  n = 35
)

sd_40 <- tibble(
  team_sd = quantiles,
  sd = sds_40,
  min = sapply(sds_40, function(.x) inverseCDF(0.025, phcauchy, .x)),
  max = sapply(sds_40, function(.x) inverseCDF(0.975, phcauchy, .x)),
  width = max - min,
  n = 40
)

bind_rows(sd_20, sd_25,sd_30, sd_35, sd_40) %>%
  ggplot(aes(team_sd, width, colour = as.factor(n))) +
  geom_line() +
  geom_point(size = 4) +
  geom_hline(yintercept = 1) +
  annotate("text", 0.9, 1.3, label = "1 SD") +
  geom_hline(yintercept = 2, linetype = "dashed") +
  annotate("text", 0.9, 2.3, label = "2 SD") +
  geom_hline(yintercept = 3, linetype = "dotted") +
  annotate("text", 0.9, 3.3, label = "3 SD") +
  ylim(0, 8) +
  scale_color_brewer() +
  labs(
    title = "95% CrI width depending on teams' true SD and number of teams",
    x = "Teams' true standard deviation",
    y = "95% CrI width (standard units)",
    colour = "N of teams"
  )

ggsave("./figs/sample-size/cri-width-nteams-true-sd.pdf", width = 7, height = 5)
```