### 2022-04-19 scripts/r/teams-analyses.Rmd
✨ - prepare files with analytic questionnaire responses

---
title: "Teams analyses"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(magrittr)
library(glue)
```

```{r team-analyses}
teams_analyses <- read_csv("data/questionnaires/msa_data_analysis_questionnaire_April 19, 2022_08.06.csv") %>%
  # Get rid of unused columns
  select(-(EndDate:UserLanguage))

# Extract the full questions text for later use
full_q_text <- teams_analyses[1,]

teams_analyses %<>% slice(-(1:2)) %>%
  # Remove test submissions
  filter(
    str_length(Q2) > 1,
    Q7 != "tba",
    StartDate != "2022-04-06 09:41:44"
  ) %>%
  # Clean
  mutate(
    Q17 = str_remove_all(Q17, '"'),
    Q17 = str_replace(Q17, "speech-and-recreation", "procambarus_mahogoni"),
    Q17 = str_replace(Q17, "paralichtys_undulatus", "paralichthys_undulatus")
  )

# Include missing OSF component names, due to inclusion of question after a few first submissions.
teams_analyses$Q17[1] <- "lasionycteris_altavela"
teams_analyses$Q16[1] <- "https://osf.io/aes8n/?view_only=232b2de4e8254af5832d63baa2a9450a"
teams_analyses$Q17[2] <- "petauroides_fistulator"
teams_analyses$Q16[2] <- "https://osf.io/k5bxr/?view_only=8d16018b339f4db58584bec4a89ccd0e"
teams_analyses$Q17[3] <- "pervagor_meeki"
teams_analyses$Q16[3] <- "https://osf.io/sfhn5/?view_only=a8cec2e3e26043f2ab0218c7a02694f2"

write_csv(teams_analyses, "data/questionnaires/msa_data_analysis_questionnaire_cleaned.csv")
```

```{r review, eval=FALSE}
the_sheet <- "https://docs.google.com/spreadsheets/d/1vFPjZkB4naCNj10v0BuEWQDvo34okmcN1Mv7VkZL9po/edit#gid=0"

review <- read_sheet(the_sheet) %>%
  left_join(teams_analyses %>% select(Q17, Q16), by = c("Animal" = "Q17"))
```

```{r}
dir.create("data/questionnaires/msa_data_analysis_files/", showWarnings = FALSE)

n_teams <- nrow(teams_analyses)

for (team in 1:n_teams) {
  animal <- teams_analyses$Q17[team]
  file_name <- as.character(glue("data/questionnaires/msa_data_analysis_files/{animal}.txt"))

  txt <- ""

  for (q in 4:15) {
    txt <- paste(
      txt,
      full_q_text[1,q], "\n",
      teams_analyses[team,q],
      "\n\n",
      "------------------------------------------------------------------------"
      , "\n\n",
      sep = "\n"
    )
  }

  write_file(txt, file_name)
}
```