### 2022-04-21 scripts/r/03-analyses-coding.Rmd
✨ - new analysis scripts

---
title: "03 - Analyses coding"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(ggmosaic)
library(googlesheets4)
library(magrittr)
library(glue)
```

```{r coding}
the_sheet <- "https://docs.google.com/spreadsheets/d/1OOXB-8Uk-fh_urw0Lm4DC0jBueXLIb4Bm-a9--UVXOw/edit?usp=sharing"

coding <- read_sheet(the_sheet) %>%
  mutate(
    outcome = factor(outcome, levels = c("duration", "f0", "formants", "intensity", "harmonics", "SNR", "multivariate", "other")),
    temporal_window = factor(temporal_window, levels = c("segment", "syllable", "word", "phrase", "sentence", "other"))
  )

```

```{r}
coding %>%
  ggplot(aes(model)) +
  geom_bar()
```

```{r}
coding %>%
  drop_na(outcome, temporal_window) %>%
  ggplot() +
  geom_mosaic(aes(x = product(outcome, temporal_window), fill = outcome))
```

```{r}
coding %>%
  drop_na(outcome, temporal_window) %>%
  filter(
    outcome %in% c("duration", "f0", "intensity"),
    temporal_window %in% c("segment", "word", "phrase", "sentence")
  ) %>%
  droplevels() %>%
  ggplot() +
  geom_mosaic(aes(x = product(outcome, temporal_window), fill = outcome))
```