### 2022-06-10 scripts/r/03-analyses-coding.Rmd
✨ - add code to download analyses

---
title: "03 - Analyses coding"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(ggmosaic)
library(googlesheets4)
library(magrittr)
library(glue)
library(osfr)
library(cli)
```

# Read coding sheet

```{r coding}
the_sheet <- "https://docs.google.com/spreadsheets/d/1OOXB-8Uk-fh_urw0Lm4DC0jBueXLIb4Bm-a9--UVXOw/edit?usp=sharing"

coding <- read_sheet(the_sheet) %>%
  mutate(
    outcome = factor(outcome, levels = c("duration", "f0", "formants", "intensity", "harmonics", "SNR", "multivariate", "other")),
    temporal_window = factor(temporal_window, levels = c("segment", "syllable", "word", "phrase", "sentence", "other"))
  )

```

# Download OSF components

First, we get a list of the OSF components in the Teams Analyses component. (It takes a few seconds)

```{r}
assigned <- coding %>%
  select(team, animal, assigned_to) %>%
  distinct()

comps <- osf_retrieve_node("https://osf.io/n3fyd/") %>%
  osf_ls_nodes(n_max = Inf) %>%
  left_join(y = assigned, by = c("name" = "animal"))
```

Now we get only those components that are assigned to you. Tell me who you are when prompt.

```{r}
who <- menu(c("Ste", "Joseph", "Timo"), "Who are you?")

whos <- c("sc", "jvc", "tr")
your_comps <- comps %>%
  filter(assigned_to == whos[who])
dir.create(glue("./data/analyses/{whos[who]}"), showWarnings = FALSE, recursive = TRUE)
```

And now we download the components in `data/analyses/`

```{r}
for (comp in 1:nrow(your_comps)) {
  name <- your_comps$name[comp]
  files <- osf_ls_files(your_comps[comp,])
  dir <- glue("./data/analyses/{whos[who]}/{name}")
  dir.create(dir, showWarnings = FALSE, recursive = TRUE)
  cli_h1(name)
  osf_download(
    files,
    path = dir,
    recurse = TRUE,
    conflicts = "skip",
    verbose = TRUE, progress = TRUE
  )
}
```

# Basic plots

```{r}
coding %>%
  ggplot(aes(model)) +
  geom_bar()
```

```{r}
coding %>%
  drop_na(outcome, temporal_window) %>%
  ggplot() +
  geom_mosaic(aes(x = product(outcome, temporal_window), fill = outcome))
```

```{r}
coding %>%
  drop_na(outcome, temporal_window) %>%
  filter(
    outcome %in% c("duration", "f0", "intensity"),
    temporal_window %in% c("segment", "word", "phrase", "sentence")
  ) %>%
  droplevels() %>%
  ggplot() +
  geom_mosaic(aes(x = product(outcome, temporal_window), fill = outcome))
```