### 2022-06-11 scripts/r/04-refit/refit-jvc.Rmd
push refit-jvc for testing

---
title: "Re-fit"
subtitle: "JVC"
output: 
  html_document:
    toc: true
date: "Last update: `r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE
  )
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
#| label: libs
library("here")
library("stringr")
library("dplyr")
library("readr")
library("lme4")
library("brms")
```

# anomalocaris_ornata
# anthracoceros_coronata
# arapaima_modularis
# comanthina_maculatus
# cromileptes_saxatilis
# ctenosaura_limax
# gnathosaurus_canadensis
# gymnothorax_spinulosus
# haematopus_fossor

# lasionycteris_altavela

```{r}
#| label: lasionycteris_altavela-load-data
# Relevant code is in 'analysis.Rmd'
# Starting at line 327

intensity_final <- read_csv(
  here("data", "analyses", "jvc", "lasionycteris_altavela", 
    "intensity_df.csv")) # load this is in

# just look at noun focus trials
intensity_DF <- subset(intensity_final, label == "NF")

#
# JVC: ORIGINAL MODEL (singular fit)
#
intensity_model_lmer <- lmer(Intensity ~ typ_mean + (1|target_name)+ (1|filename) + (1|filename:target_name),
                       data = intensity_DF)

#
# JVC: Standardized model
# - Model contains two variables to standardize: Intensity and typ_mean 
# - Both variables are continuous, parameter estimate for typ_mean will be 
#   effect size (named effect_con)
#
MSA_lasionycteris_altavela_intensity_DF <- intensity_DF %>% 
  mutate(int_z = (Intensity - mean(Intensity)) / sd(Intensity), 
    effect_con = (typ_mean - mean(typ_mean)) / sd(typ_mean))

lasionycteris_altavela_1_int_con <- brm(
  int_z ~ effect_con + 
    (1|target_name) + (1|filename) + (1|filename:target_name), 
  data = MSA_lasionycteris_altavela_intensity_DF, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "lasionycteris_altavela_1_int_con")
  )
```

# linckia_nattereri
# lycodes_bradfieldi
# naso_cassivellaunos
# trapezia_cantonensis
# trigonias_lachneri