### 2022-06-14 scripts/r/04-refit/refit-jvc.Rmd
add haematopus_fossor and test

---
title: "Re-fit"
subtitle: "JVC"
output: 
  html_document:
    toc: true
date: "Last update: `r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE
  )
knitr::opts_knit$set(root.dir = here::here())
```

```{r}
#| label: libs
library("here")
library("fs")
library("dplyr")
library("tidyr")
library("readr")
library("forcats")
library("ggplot2")
library("stringr")
library("strex")
library("purrr")
library("lme4")
library("brms")
library("broom")
library("broom.mixed")
```

```{r}
#| label: helpers
#| include: false
# Plotting theme because Im extra
clean_theme <- function(...) {
  list(
    theme_minimal(), 
    theme(
      axis.title.y = element_text(size = rel(.9), hjust = 0.95),
      axis.title.x = element_text(size = rel(.9), hjust = 0.95),
      panel.grid.major = element_line(colour = 'grey90', size = 0.15),
      panel.grid.minor = element_line(colour = 'grey90', size = 0.15))
  )
}

# Functional sequence to get relevant info from Rds files
clean_up <- . %>% 
  tidy(effects = "fixed") %>% 
  suppressWarnings() %>% 
  filter(term %in% c("effect_cattypical", "effect_con")) %>% 
  select(term, estimate, se = std.error)
```

# Refits

## anomalocaris_ornata

## anthracoceros_coronata

## arapaima_modularis

## comanthina_maculatus

## cromileptes_saxatilis

## ctenosaura_limax

## gnathosaurus_canadensis

## gymnothorax_spinulosus

## haematopus_fossor

```{r}
#| label: haematopus_fossor

directory <- here("data", "trial-lists")
list_of_files = dir(here("data", "trial-lists"))

trial_data = data.frame()
for(file in list_of_files) {
  this_file = read.csv(paste0(directory, "/", file), stringsAsFactors = FALSE)
  trial_data = rbind(trial_data, this_file)
}

trial_data$typicality = factor(trial_data$typicality)
contrasts(trial_data$typicality) = contr.helmert(3)
dimnames(contrasts(trial_data$typicality))[[2]] = c("atypical", "typical")

#
# DURATION
#

# Get duration data
duration <- read_csv(here("data", "analyses", "jvc", "haematopus_fossor", 
  "details", "analysis", "wordDurations.csv")) %>% 
  filter(adj != "")

# Merge in trial data
duration$speaker = gsub(".TextGrid", "", duration$participant)
duration_full = merge(duration, trial_data)

# Remove errors
duration_full = filter(duration_full, comments!="hesitation break")

# Centre relevant variables
duration_full$adjdur.c = scale(duration_full$adjdur, scale=FALSE)
duration_full$noundur.c = scale(duration_full$noundur, scale=FALSE)
duration_full$framedur.c = scale(duration_full$framedur, scale=FALSE)

# LME models
# Adjective duration
adj_duration_model = lmer(adjdur.c ~ trial * typicality + (1 | adj) + (1 | speaker), duration_full)

# Noun duration
noun_duration_model = lmer(noundur.c ~ trial * typicality + (1 | noun) + (1 | speaker), duration_full)

# Frame duration
frame_duration_model = lmer(framedur.c ~ trial * typicality + (1 | speaker), duration_full)

#
# JVC: setup df
#
MSA_duration_df <- duration_full %>% 
  mutate(across(c("adjdur", "noundur", "framedur"), .fns = scale, 
    .names = "{.col}_z")) %>% 
  mutate(effect_cat = C(typicality, treatment))

# Adjective duration
haematopus_fossor_1_adjdur_cat <- brm(
  adjdur_z ~ trial * effect_cat + 
    (1 | adj) + (1 | speaker), 
  data = MSA_duration_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_1_adjdur_cat")
  )

# Noun duration
haematopus_fossor_2_noundur_cat <- brm(
  noundur_z ~ trial * effect_cat + 
    (1 | noun) + (1 | speaker), 
  data = MSA_duration_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_2_noundur_cat")
)

# Frame duration
haematopus_fossor_3_framedur_cat <- brm(
  framedur_z ~ trial * effect_cat + 
    (1 | speaker), 
  data = MSA_duration_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_3_framedur_cat")
)

#
# JVC: same for Vowel
#
vowel <- read_csv(here("data", "analyses", "jvc", "haematopus_fossor", 
  "details", "analysis", "vowelMeasurements.csv"), na = "--undefined--") %>% 
  # Get rid of data with no word or vowel label
  filter(word != "", vowel != "")

# Merge in trial data
vowel$speaker = gsub("\\.", "", vowel$participant)
vowel_full = merge(vowel, trial_data)

# Fix typos in word names
vowel_full$word = gsub("banana","banane", vowel_full$word)
vowel_full$word = gsub("braunen0","braunen", vowel_full$word)
vowel_full$word = gsub("erbeere","erdbeere", vowel_full$word)
vowel_full$word = gsub("greunen","gruenen", vowel_full$word)
vowel_full$word = gsub("gruenen ","gruenen", vowel_full$word)
vowel_full$word = gsub("orangen ","orangen", vowel_full$word)
vowel_full$word = gsub("orangenen","orangen", vowel_full$word)
vowel_full$word = gsub("Paprika","paprika", vowel_full$word)
vowel_full$word = gsub("traubenn","trauben", vowel_full$word)

# Fix typos in vowel label and exclude erroneously segmented segments
vowel_full = vowel_full %>% filter(!vowel %in% c(" ", "@0", "&0", "a0", "e0", "r"))
vowel_full$vowel = gsub("gelben","E1", vowel_full$vowel)
vowel_full$vowel = gsub("E1o","E1", vowel_full$vowel)
vowel_full$vowel = gsub("01","o1", vowel_full$vowel)
vowel_full$vowel = gsub("I.*","I1", vowel_full$vowel)
vowel_full$vowel = gsub("y.*","y1", vowel_full$vowel)

# Define nouns and adjectives
list_of_nouns = c("aprikose", "banane", "bohnen", "erbsen", "erdberre", "gurken", "kartoffeln", "kirche", "mandarine", "moehre", "paprika", "tomate", "trauben", "walnuss", "zitrone")
vowel_full$pos = ifelse(vowel_full$word %in% list_of_nouns, "noun", "adj")

# Define centralisation metric
vowel_full = vowel_full %>% 
  group_by(speaker) %>% 
  mutate(centralisation = sqrt((F1 - mean(F1, na.rm=TRUE))^2 + (F2 - mean(F2, na.rm=TRUE))^2))

#############
# Analyses
#############

adj_full = vowel_full %>% filter(pos == "adj")
noun_full = vowel_full %>% filter(pos == "noun")

# Centre the variables
noun_full$duration.c = scale(noun_full$duration, scale=FALSE)
noun_full$intensity.c = scale(noun_full$intensity, scale=FALSE)
noun_full$centralisation.c = scale(noun_full$centralisation, scale=FALSE)
adj_full$duration.c = scale(adj_full$duration, scale=FALSE)
adj_full$intensity.c = scale(adj_full$intensity, scale=FALSE)
adj_full$centralisation.c = scale(adj_full$centralisation, scale=FALSE)

# Duration analysis
adj_vowel_duration_model = lmer(duration.c ~ trial * typicality + (1 | word) + (1 | speaker), adj_full)
noun_vowel_duration_model = lmer(duration.c ~ trial * typicality + (1 | word) + (1 | speaker), noun_full)

# Intensity analysis
adj_vowel_int_model = lmer(intensity.c ~ trial * typicality + (1 | word) + (1 | speaker), adj_full)
noun_vowel_int_model = lmer(intensity.c ~ trial * typicality + (1 | word) + (1 | speaker), noun_full)

# Centralisation analysis
adj_vowel_cent_model = lmer(centralisation.c ~ trial * typicality + (1 | word) + (1 | speaker), adj_full)
noun_vowel_cent_model = lmer(centralisation.c ~ trial * typicality + (1 | word) + (1 | speaker), noun_full)

#
# JVC: standardize predictors
#
MSA_vowel_noun_df <- noun_full %>% 
  mutate(across(c("duration", "intensity", "centralisation"), .fns = scale, 
    .names = "{.col}_z")) %>% 
  mutate(effect_cat = C(typicality, treatment))

MSA_vowel_adj_df <- adj_full %>% 
  mutate(across(c("duration", "intensity", "centralisation"), .fns = scale, 
    .names = "{.col}_z")) %>% 
  mutate(effect_cat = C(typicality, treatment))

#
# JVC: refit models
#

# Duration analysis
haematopus_fossor_4_adjvoweldur_cat <- brm(
  duration_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_adj_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_4_adjvoweldur_cat")
)

haematopus_fossor_5_nounvoweldur_cat <- brm(
  duration_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_noun_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_5_nounvoweldur_cat")
)

# Intensity analysis
haematopus_fossor_6_adjvowelint_cat <- brm(
  intensity_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_adj_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_6_adjvowelint_cat")
)

haematopus_fossor_7_nounvowelint_cat <- brm(
  intensity_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_noun_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_7_nounvowelint_cat")
)

# Centralisation analysis
haematopus_fossor_8_adjvowelcent_cat <- brm(
  centralisation_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_adj_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_8_adjvowelcent_cat")
)

haematopus_fossor_9_nounvowelcent_cat <- brm(
  centralisation_z ~ trial * effect_cat + 
    (1 | word) + (1 | speaker), 
  data = MSA_vowel_noun_df, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "haematopus_fossor_9_nounvowelcent_cat")
)
```

## lasionycteris_altavela

```{r}
#| label: lasionycteris_altavela
# Relevant code is in 'analysis.Rmd'
# Starting at line 327

intensity_final <- read_csv(
  here("data", "analyses", "jvc", "lasionycteris_altavela", 
    "intensity_df.csv")) # load this is in

# just look at noun focus trials
intensity_DF <- subset(intensity_final, label == "NF")

#
# JVC: ORIGINAL MODEL (singular fit)
#
intensity_model_lmer <- lmer(Intensity ~ typ_mean + (1|target_name)+ (1|filename) + (1|filename:target_name),
                       data = intensity_DF)

#
# JVC: Standardized model
# - Model contains two variables to standardize: Intensity and typ_mean 
# - Both variables are continuous, parameter estimate for typ_mean will be 
#   effect size (named effect_con)
#
MSA_lasionycteris_altavela_intensity_DF <- intensity_DF %>% 
  mutate(int_z = (Intensity - mean(Intensity)) / sd(Intensity), 
    effect_con = (typ_mean - mean(typ_mean)) / sd(typ_mean))

lasionycteris_altavela_1_int_con <- brm(
  int_z ~ effect_con + 
    (1|target_name) + (1|filename) + (1|filename:target_name), 
  data = MSA_lasionycteris_altavela_intensity_DF, 
  cores = 4, threads = threading(2, grainsize = 100), 
  backend = "cmdstanr", 
  file = here("data", "analyses", "models", "lasionycteris_altavela_1_int_con")
  )
```

## linckia_nattereri

## lycodes_bradfieldi

## naso_cassivellaunos

## trapezia_cantonensis

## trigonias_lachneri

# Results

## Load models

```{r}
#| label: load-models

# This code will need to change eventually to pull rds files from OSF... for now it assumes they are stored in a folder called "models"
msa_models <- dir_ls(path = here("data", "analyses", "models"), regexp = ".rds$") %>% 
  as_tibble() %>% 
  transmute(path = value, 
    mod_name = str_remove(path, here("data", "analyses", "models/")), 
    mod_name = str_remove(mod_name, ".rds"), 
    model = map(path, ~ readRDS(file = .))) %>% 
  separate(mod_name, 
    into = c("word1", "word2", "mod_n", "outcome", "typicality"), 
    sep = "_", remove = F) %>% 
  unite(group, word1, word2, sep = "_") %>% 
  select(group, mod_name:model) %>% 
  mutate(sum = map(model, clean_up)) %>% 
  suppressWarnings() # because it warns every time that some coef have underscores and its annoying
```

## Table

```{r}
#| label: effect-table

# Simple pandoc table of the models, estimates, se, etc.
msa_models %>% 
  unnest(sum) %>% 
  select(-model) %>% 
  knitr::kable(format = "pandoc")

```

## Plot

```{r}
#| label: forest-plot

# Simple forest plot of the estimates +/- SE
msa_models %>% 
  unnest(sum) %>% 
  mutate(typicality_lab = if_else(typicality == "cat", 
    "Typicality = categorical", "Typicality = continuous"), 
    mod_name = fct_reorder(mod_name, estimate)) %>% 
  ggplot() + 
  facet_wrap(~ typicality_lab, ncol = 1, scales = "free") + 
  aes(x = estimate, y = mod_name) + 
  geom_segment(aes(x = estimate - se, xend = estimate + se, yend = mod_name), 
    color = "#cc0033", alpha = 0.5, size = 3, lineend = "round") + 
  geom_point(color = "#cc0033", pch = 21, stroke = 3, size = 3.1) + 
  labs(y = "Model", x = "Estimate ± SE") + 
  clean_theme() + 
  theme(strip.text.x = element_text(hjust = 0))
```