### 2022-06-16 data/osf/norming/README.Rmd

---
title: "Norming study"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir = here::here())
library(tidyverse)
library(osfr)
```

Last update: `r Sys.Date()`

# Overview

This directory contains the files of the norming study, designed to obtain typicality ratings of re-colored images of everyday objects.

There are 2 subdirectories:

- The `images/` directory contains the re-colored images used in the norming study.
- The `ratings/` directory holds an aggregated `ratings.csv` file with the typicality ratings of each image/object for each participant.

# Ratings

The `ratings.csv` file represents the raw data collected from the norming experiment and has the following columns:

- `subject_id`: The participant/trial id.
- `object`: The name of the object illustrated in the image.
- `typicality`: The typicality rating (numeric, `1-100`, `1` = most atypical, `100` = most typical).
- `colour`: The color of the object in the image.
- `object_en`: The english translation of `object`
- `colour_en`: The english translation of `colour`

The `ratings_summary.csv` represents the aggregated typicality ratings across combinations of object and color.
The file has the following columns:

- `object`: The name of the object illustrated in the image.
- `colour`: The color of the object in the image.
- `object_en`: The english translation of `object`.
- `colour_en`: The english translation of `colour`.
- `typ_mean`: The mean typicality rating.
- `typ_sd`: The standard deviation of the typicality ratings .
- `typ_median`: The median typicality rating.