### 2022-09-09 scripts/r/07_team-sds.Rmd
join df with relevant data from vars from google sheet

---
title: "Extract team means and SDs"
author: "Joseph V. Casillas"
output: 
  html_document: 
    toc: yes
    toc_depth: 1
    number_sections: yes
date: "Last update: `r Sys.Date()`"
---

```{r}
#| label: setup
#| include: false

# Knitr options
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE
  )

knitr::opts_knit$set(root.dir = here::here())

# Load libs
library("here")
library("tidyr")
library("dplyr")
library("osfr")
library("speakr")
library("glue")
library("googlesheets4")
library("forcats")
library("readr")
library("readxl")
library("brms")
library("lme4")
library("purrr")

```

# Overview

The purpose of this script is to extract the mean and standard deviation of every dependent variable (if it is f0, duration, or intensity) from each team included in the meta-analysis (`05_meta-analyses.Rmd`). 
Most of the code needed to accomplish this is already written, thus the script recycles already completed work and, where necessary, we use a few helpers to reduce work. 
The workflow is the following:

1. Get a list of the subset of teams that need to be downloaded (Max = 30 because the final model only includes 30 teams). 
2. Download the OSF components of the aforementioned subset of teams.
3. Use the code found in `04_refit` (i.e., `refit_jvc.Rmd`, `refit_sc.Rmd` and `refit_tr.Rmd`) to get the final data frames used in the original analyses. 
4. Calculate mean/SD of the raw DVs (only f0, duration, and intensity). 
5. Store means/SDs in a separate .csv file in `data/analyses`.

# List of teams

```{r}
#| label: get-teams
#| message: false
#| warning: false

# Get team names from final meta-analytic model
teams <- readRDS(here("./data/analyses/msa_models.rds")) %>% 
  pull(animal) %>% 
  unique

# Get coding excel from google drive
assigned <- read_sheet("https://docs.google.com/spreadsheets/d/1OOXB-8Uk-fh_urw0Lm4DC0jBueXLIb4Bm-a9--UVXOw/edit?usp=sharing")

# Use teams to subset excel file with all analyses, get those that include f0, 
# duration, and intensity
final_teams <- assigned %>% 
  filter(animal %in% teams, outcome %in% c("f0", "duration", "intensity"), 
         exclude == "no")

# Just the names
final_team_names <- final_teams %>% 
  pull(animal) %>% 
  unique

# Create dataframe (lookup table) with final teams names and the assigned coder 
final_team_name_and_coder <- final_teams %>% 
  select(animal, assigned_to) %>% 
  distinct()
```

Of the 30 teams included in the meta-analysis, `r length(final_team_names)` included either F0, duration or intensity in the analyses. 
Next, we download the OSF components of those `r length(final_team_names)` teams.

# Download OSF team components

```{r}
#| label: download-setup
#| eval: false

# Get OSF team components
osf_teams <- osf_retrieve_node("n3fyd")
teams_comps <- osf_ls_nodes(osf_teams, n_max = Inf)

# Function to download team components
#  - only those in final analysis
#  - only those including F0, pitch, duration
#  - separate by coder and put components in coder-specific directories
download_components <- function(coder) {
  # Get vector of team names by coder assignment
  my_assignment <- filter(final_team_name_and_coder, assigned_to == coder) %>% 
    pull(animal)

  # Loop through teams in my_assignment and download the corresponding OSF comp
  # to /data/analyses/{coder}
  for (team in 1:length(my_assignment)) {
    skip_to_next <- FALSE

    name <- my_assignment[team]
    dir.create(glue("./data/analyses/{coder}/{name}"), 
               recursive = TRUE, showWarnings = FALSE)

    tryCatch(
      osf_ls_files(teams_comps[team,], type = "folder") %>%
        osf_download(
          path = glue("./data/analyses/{coder}/{name}"),
            recurse = TRUE, verbose = TRUE, progress = TRUE, conflicts = "skip"
          ),
      error = function(e) {skip_to_next <<- TRUE}
    )

    if(skip_to_next) { next }
  }
}
```

```{r}
#| label: download-components
#| eval: false

# Download JVC
download_components(coder = "jvc")

# Download SC
download_components(coder = "sc")

# Download TR
download_components(coder = "tr")
```

# Get means and SDs

In order to get the data frame from each teams analysis, we will recycle code from `04_refit` (i.e., `refit_jvc.Rmd`, `refit_sc.Rmd` and `refit_tr.Rmd`). 
We can pull in specific lines of code with the `source_part` function. 
If we can't source the lines directly without touching the refit_xx.Rmd files (because *some* people refuse to use `here`), then we will have to copy/paste code to this file and adjusts the paths like heathens.

```{r}
#| label: source-part
source_part <- function(file, start, end, ...) {
  # Set lines
  file.lines <- scan(
    file, 
    what = character(), 
    skip = start - 1, 
    nlines = end - start + 1, 
    sep = '\n'
    )
  # Collapse lines
  file.lines.collapsed <- paste(file.lines, collapse = '\n')
  # Source selection
  source(textConnection(file.lines.collapsed), ...)
}
```

We have the following number of OSF components for each coder:

```{r}
#| label: breakdown-by-coder
final_team_name_and_coder %>% 
  group_by(assigned_to) %>% 
  summarize(n = n_distinct(animal)) %>% 
  knitr::kable()
```

## JVC

```{r}
#| label: jvc-teams
#| echo: false
#| results: 'asis'
final_team_name_and_coder %>% 
  filter(assigned_to == "jvc") %>% 
  select(animal) %>% 
  arrange(animal) %>% 
  mutate(animal = glue("- {animal}")) %>% 
  pull

```

### anomalocaris_ornata

```{r}
#| label: anomalocaris_ornata
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 131, 
  end = 181
  )

# Means and SDs
anomalocaris_ornata_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "anomalocaris_ornata", 1, "intensity", mean(df.int$Mean_dB),        sd(df.int$Mean_dB), 
  "anomalocaris_ornata", 2, "f0",        mean(under$f0sem),             sd(under$f0sem), 
  "anomalocaris_ornata", 3, "f0",        mean(over$f0sem),              sd(over$f0sem), 
  "anomalocaris_ornata", 4, "duration",  mean(vowel.df$Duration_in_ms), sd(vowel.df$Duration_in_ms)
)

# Check n models == n DVs
final_teams %>% filter(animal == "anomalocaris_ornata") %>% nrow == nrow(anomalocaris_ornata_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "anomalocaris_ornata") %>% pull(outcome) == anomalocaris_ornata_df %>% pull(dv)
```

### anthracoceros_coronata

```{r}
#| label: anthracoceros_coronata
#| eval: true

data_long <- read_csv(here("data", "analyses", "jvc", "anthracoceros_coronata", 
  "data", "data_from_tgs.csv")) %>% 
  mutate(A_dur_log = log(A_end - A_start), 
         N_dur_log = log(N_end - N_start))

# Means and SDs
anthracoceros_coronata_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "anthracoceros_coronata", 1, "f0",        mean(data_long$A_max_pitch),  sd(data_long$A_max_pitch), 
  "anthracoceros_coronata", 2, "f0",        mean(data_long$N_max_pitch, na.rm = T),  sd(data_long$N_max_pitch, na.rm = T), 
  "anthracoceros_coronata", 3, "intensity", mean(data_long$A_max_intens), sd(data_long$A_max_intens), 
  "anthracoceros_coronata", 4, "intensity", mean(data_long$N_max_intens), sd(data_long$N_max_intens), 
  "anthracoceros_coronata", 5, "duration",  mean(data_long$A_dur_log),    sd(data_long$A_dur_log), 
  "anthracoceros_coronata", 6, "duration",  mean(data_long$N_dur_log),    sd(data_long$N_dur_log), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "anthracoceros_coronata") %>% nrow == nrow(anthracoceros_coronata_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "anthracoceros_coronata") %>% pull(outcome) == anthracoceros_coronata_df %>% pull(dv)
```

### arapaima_modularis

```{r}
#| label: arapaima_modularis
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 392, 
  end = 441
  )

# Means and SDs
arapaima_modularis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "arapaima_modularis", 1, "f0", mean(data_for_analysis_noOL$mean_pitch), sd(data_for_analysis_noOL$mean_pitch)
)

# Check n models == n DVs
final_teams %>% filter(animal == "arapaima_modularis") %>% nrow == nrow(arapaima_modularis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "arapaima_modularis") %>% pull(outcome) == arapaima_modularis_df %>% pull(dv)
```

### clione_dorsalis

```{r}
#| label: clione_dorsalis
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 1379, 
  end = 1395
  )

# Means and SDs
clione_dorsalis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "clione_dorsalis", 1, "duration", mean(data_formant_midpoint$word.duration), sd(data_formant_midpoint$word.duration)
)

# Check n models == n DVs
final_teams %>% filter(animal == "clione_dorsalis") %>% nrow == nrow(clione_dorsalis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "clione_dorsalis") %>% pull(outcome) == clione_dorsalis_df %>% pull(dv)
```

### comanthina_maculatus

```{r}
#| label: comanthina_maculatus
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 478, 
  end = 489
  )

# Means and SDs
comanthina_maculatus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "comanthina_maculatus", 1, "duration", mean(MSA_comanthina_maculatus_df$tonic_vowel_duration), sd(MSA_comanthina_maculatus_df$tonic_vowel_duration), 
  "comanthina_maculatus", 2, "intensity", mean(MSA_comanthina_maculatus_df$tonic_vowel_intensity), sd(MSA_comanthina_maculatus_df$tonic_vowel_intensity),  
  "comanthina_maculatus", 3, "f0", mean(MSA_comanthina_maculatus_df$tonic_vowel_F0), sd(MSA_comanthina_maculatus_df$tonic_vowel_F0), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "comanthina_maculatus") %>% nrow == nrow(comanthina_maculatus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "comanthina_maculatus") %>% pull(outcome) == comanthina_maculatus_df %>% pull(dv)
```

### cromileptes_saxatilis

```{r}
#| label: cromileptes_saxatilis
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 527, 
  end = 530
  )

# Means and SDs
cromileptes_saxatilis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "cromileptes_saxatilis", 1, "duration", mean(cromileptes_saxatilis_durs$colour_dur), sd(cromileptes_saxatilis_durs$colour_dur)
)

# Check n models == n DVs
final_teams %>% filter(animal == "cromileptes_saxatilis") %>% nrow == nrow(cromileptes_saxatilis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "cromileptes_saxatilis") %>% pull(outcome) == cromileptes_saxatilis_df %>% pull(dv)
```

### gnathosaurus_canadensis

```{r}
#| label: gnathosaurus_canadensis
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 602, 
  end = 606
  )

# Means and SDs
gnathosaurus_canadensis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "gnathosaurus_canadensis", 1, "f0", mean(gnathosaurus_canadensis_typ$median_st_diff), sd(gnathosaurus_canadensis_typ$median_st_diff)
)

# Check n models == n DVs
final_teams %>% filter(animal == "gnathosaurus_canadensis") %>% nrow == nrow(gnathosaurus_canadensis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "gnathosaurus_canadensis") %>% pull(outcome) == gnathosaurus_canadensis_df %>% pull(dv)
```

### gymnothorax_spinulosus

```{r}
#| label: gymnothorax_spinulosus
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 644, 
  end = 662
  )

# Means and SDs
gymnothorax_spinulosus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "gymnothorax_spinulosus", 1, "intensity", mean(gymnothorax_spinulosus_int$measure), sd(gymnothorax_spinulosus_int$measure)
)

# Check n models == n DVs
final_teams %>% filter(animal == "gymnothorax_spinulosus") %>% nrow == nrow(gymnothorax_spinulosus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "gymnothorax_spinulosus") %>% pull(outcome) == gymnothorax_spinulosus_df %>% pull(dv)
```

### haematopus_fossor

```{r}
#| label: haematopus_fossor
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 714, 
  end = 845
  )

# Means and SDs
haematopus_fossor_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "haematopus_fossor", 1, "duration", mean(duration_full$framedur), sd(duration_full$framedur), 
  "haematopus_fossor", 2, "duration", mean(duration_full$adjdur), sd(duration_full$adjdur), 
  "haematopus_fossor", 3, "duration", mean(duration_full$noundur), sd(duration_full$noundur), 
  "haematopus_fossor", 4, "duration", mean(adj_full$duration), sd(adj_full$duration), 
  "haematopus_fossor", 5, "duration", mean(noun_full$duration), sd(noun_full$duration), 
  "haematopus_fossor", 6, "intensity", mean(adj_full$intensity), sd(adj_full$intensity), 
  "haematopus_fossor", 7, "intensity", mean(noun_full$intensity), sd(noun_full$intensity), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "haematopus_fossor") %>% nrow == nrow(haematopus_fossor_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "haematopus_fossor") %>% pull(outcome) == haematopus_fossor_df %>% pull(dv)
```

### lasionycteris_altavela

```{r}
#| label: lasionycteris_altavela
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 955, 
  end = 962
  )

# Means and SDs
lasionycteris_altavela_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "lasionycteris_altavela", 1, "intensity", mean(intensity_DF$Intensity), sd(intensity_DF$Intensity),
)

# Check n models == n DVs
final_teams %>% filter(animal == "lasionycteris_altavela") %>% nrow == nrow(lasionycteris_altavela_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "lasionycteris_altavela") %>% pull(outcome) == lasionycteris_altavela_df %>% pull(dv)
```

### naso_cassivellaunos

```{r}
#| label: naso_cassivellaunos
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 1054, 
  end = 1067
  )

# Means and SDs
naso_cassivellaunos_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "naso_cassivellaunos", 2, "f0", mean(msa$DeltaF0_rel), sd(msa$DeltaF0_rel), 
  "naso_cassivellaunos", 3, "f0", mean(msa$sync_rel), sd(msa$sync_rel),
)

# Check n models == n DVs
final_teams %>% filter(animal == "naso_cassivellaunos") %>% nrow == nrow(naso_cassivellaunos_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "naso_cassivellaunos") %>% pull(outcome) == naso_cassivellaunos_df %>% pull(dv)
```

### psittacula_scabriculus

```{r}
#| label: psittacula_scabriculus
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 1295, 
  end = 1327
  )

# Means and SDs
psittacula_scabriculus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "psittacula_scabriculus", 1, "duration", mean(MSAcomplete$avgdur), sd(MSAcomplete$avgdur), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "psittacula_scabriculus") %>% nrow == nrow(psittacula_scabriculus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "psittacula_scabriculus") %>% pull(outcome) == psittacula_scabriculus_df %>% pull(dv)
```

### trapezia_cantonensis

```{r}
#| label: trapezia_cantonensis
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_jvc.Rmd"), 
  start = 1206, 
  end = 1214
  )

# Means and SDs
trapezia_cantonensis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "trapezia_cantonensis", 4, "duration", mean(MSA_trapezia_cantonensis_data$dur), sd(MSA_trapezia_cantonensis_data$dur), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "trapezia_cantonensis") %>% nrow == nrow(trapezia_cantonensis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "trapezia_cantonensis") %>% pull(outcome) == trapezia_cantonensis_df %>% pull(dv)
```

## SC

```{r}
#| label: sc-teams
#| echo: false
#| results: 'asis'
final_team_name_and_coder %>% 
  filter(assigned_to == "sc") %>% 
  select(animal) %>% 
  arrange(animal) %>% 
  mutate(animal = glue("- {animal}")) %>% 
  pull
```

### dermatolepis_aculeatus

```{r}
#| label: dermatolepis_aculeatus
#| eval: true

source_part(
 file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
 start = 21, 
 end = 32
)

msa_data <- read_csv(here("data", "analyses", "sc", "dermatolepis_aculeatus", 
  "Many Speech Analyses_measurements.csv")) %>%
  filter(exclude == "N", condition == "NF") %>%
  drop_na()

# Means and SDs
dermatolepis_aculeatus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "dermatolepis_aculeatus", 1, "f0", mean(msa_data$noun_pitch_max), sd(msa_data$noun_pitch_max), 
  "dermatolepis_aculeatus", 2, "f0", mean(msa_data$adj_pitch_max), sd(msa_data$adj_pitch_max), 
  "dermatolepis_aculeatus", 3, "f0", mean(msa_data$phrase_pitch_max), sd(msa_data$phrase_pitch_max), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "dermatolepis_aculeatus") %>% nrow == nrow(dermatolepis_aculeatus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "dermatolepis_aculeatus") %>% pull(outcome) == dermatolepis_aculeatus_df %>% pull(dv)
```

### eriphia_laterispinis

```{r}
#| label: eriphia_laterispinis
#| eval: true

# Can't source code because of path issues
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = 77, 
#  end = 88
#  )

# listing all available csv files (30 participants in the production study)
trials_files <- list.files(
    path = here("data", "osf", "production", "trial-lists/"),
    pattern = ".csv", full.names = TRUE
    )

# importing and reshaping all csv files at once
trials <- map_dfr(
    .x = trials_files,
    .f = read.csv
    )

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
msa_data <- read.csv(here("data", "analyses", "sc", "eriphia_laterispinis", 
    "processed_data", "summary_data.csv")) %>%
  filter(Condition == "NF") %>%
  left_join(x = ., y = trials, by = c("speaker", "trial") ) %>%
  select(speaker, trial, condition, mean_f0, mean_int, duration, typicality, 
    typ_mean, typ_sd, typ_median) %>%
  filter(typicality != "medium")

# Means and SDs
eriphia_laterispinis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "eriphia_laterispinis", 1, "f0",        mean(msa_data$mean_f0), sd(msa_data$mean_f0), 
  "eriphia_laterispinis", 1, "intensity", mean(msa_data$mean_int), sd(msa_data$mean_int), 
  "eriphia_laterispinis", 1, "duration",  mean(msa_data$duration), sd(msa_data$duration), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "eriphia_laterispinis") %>% nrow == nrow(eriphia_laterispinis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "eriphia_laterispinis") %>% pull(outcome) == eriphia_laterispinis_df %>% pull(dv)
```

### hoplostethus_macrosteus

```{r}
#| label: hoplostethus_macrosteus
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

msa_data <- read_csv(here("data", "analyses", "sc", "hoplostethus_macrosteus", 
    "Analysis materials", "Results_clean.csv")) %>%
  mutate(
    effect_cat = factor(typicality, levels = c("atypical", "medium", "typical")),
    color = factor(color)
  )

# Means and SDs
hoplostethus_macrosteus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "hoplostethus_macrosteus", 1, "duration",  mean(msa_data$durmsec), sd(msa_data$durmsec), 
  "hoplostethus_macrosteus", 2, "intensity", mean(msa_data$intensity_maximum), sd(msa_data$intensity_maximum), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "hoplostethus_macrosteus") %>% nrow == nrow(hoplostethus_macrosteus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "hoplostethus_macrosteus") %>% pull(outcome) == hoplostethus_macrosteus_df %>% pull(dv)
```

### neosilurus_omanensis

```{r}
#| label: neosilurus_omanensis
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

msa_data <- read_csv(here("data", "analyses", "sc", "neosilurus_omanensis", 
    "annotated_cleaned_data_v7.csv")) %>% 
  drop_na()

# Means and SDs
neosilurus_omanensis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "neosilurus_omanensis", 1, "f0", mean(msa_data$PitchRatioAtoU), sd(msa_data$PitchRatioAtoU), 
  "neosilurus_omanensis", 2, "f0", mean(msa_data$PitchRangeRatioAtoU), sd(msa_data$PitchRangeRatioAtoU), 
  "neosilurus_omanensis", 4, "f0", mean(msa_data$PitchRiseRatio), sd(msa_data$PitchRiseRatio), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "neosilurus_omanensis") %>% nrow == nrow(neosilurus_omanensis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "neosilurus_omanensis") %>% pull(outcome) == neosilurus_omanensis_df %>% pull(dv)
```

### nestor_idahoensis

```{r}
#| label: nestor_idahoensis
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

msa_data <- read_csv(here("data", "analyses", "sc", "nestor_idahoensis", 
    "msa_data_v2.csv"), na = c("", "--undefined--")) %>%
  mutate(
    f0 = log(f0) %>% as.numeric(),
    duration = log(duration) %>% as.numeric(),
    f15 = log(f15) %>% as.numeric(),
    f25 = log(f25) %>% as.numeric()
  )

# Means and SDs
nestor_idahoensis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "nestor_idahoensis", 1, "duration", mean(msa_data$duration), sd(msa_data$duration), 
  "nestor_idahoensis", 1, "f0",       mean(msa_data$f0), sd(msa_data$f0), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "nestor_idahoensis") %>% nrow == nrow(nestor_idahoensis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "nestor_idahoensis") %>% pull(outcome) == nestor_idahoensis_df %>% pull(dv)
```

### paralichthys_undulatus

```{r}
#| label: paralichthys_undulatus
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

msa_data <- read_delim(here("data", "analyses", "sc", "paralichthys_undulatus", 
    "Data files", "MSA-measurements.txt"), delim = "\t")

# Means and SDs
paralichthys_undulatus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "paralichthys_undulatus", 1, "f0", mean(msa_data$RangeST), sd(msa_data$RangeST), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "paralichthys_undulatus") %>% nrow == nrow(paralichthys_undulatus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "paralichthys_undulatus") %>% pull(outcome) == paralichthys_undulatus_df %>% pull(dv)
```

### petauroides_fistulator

```{r}
#| label: petauroides_fistulator
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

load(here("data", "analyses", "sc", "petauroides_fistulator", "mdl_data.RData"))

msa_data <- mdl_data %>% drop_na()

# Means and SDs
petauroides_fistulator_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "petauroides_fistulator", 1, "f0", mean(msa_data$pitch_st), sd(msa_data$pitch_st), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "petauroides_fistulator") %>% nrow == nrow(petauroides_fistulator_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "petauroides_fistulator") %>% pull(outcome) == petauroides_fistulator_df %>% pull(dv)
```

### saron_pictus

```{r}
#| label: saron_pictus
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

# Super long because somebody refuses to use `here` and I don't want to modify `refit_sc.Rmd`
# The following code reads in the thirty different .csv files with the trial-level data and merges them into a single dataframe
trial_dat <- list.files(path = here("data", "osf", "production", "trial-lists"),
                       pattern = "*.csv",
                       full.names = TRUE) %>%
  lapply(read_csv) %>%                              
  bind_rows

# Reading in the results file output by the Praat script
d <- read.delim(here("data", "analyses", "sc", "saron_pictus", "results.txt"),
                fileEncoding = "UTF-16")

# Saving original number of rows
N <- nrow(d)

# Getting rid of columns we won't need from trial_dat
trial_dat <- trial_dat[trial_dat$condition == "NF",]

# Changing the trial_dat colours to the inflected forms to match the utterances
trial_dat$target_colour[trial_dat$target_colour == "gelb"] <- "gelben"
trial_dat$target_colour[trial_dat$target_colour == "orange"] <- "orangen"
trial_dat$target_colour[trial_dat$target_colour == "braun"] <- "braunen"
trial_dat$target_colour[trial_dat$target_colour == "rot"] <- "roten"
trial_dat$target_colour[trial_dat$target_colour == "gruen"] <- "gruenen"

# Saving the target colours present in the data-set into a vector
colours <- c("gelben", "orangen", "braunen", "roten", "gruenen")

# Subsetting to only target nouns and adjectives
d <- d[d$word_label %in% trial_dat$target_name | d$word_label %in% trial_dat$target_colour,]
d <- droplevels(d)

# Merging the data from Praat with the trial data
d <- merge(d, trial_dat, 
            by = c("speaker", "trial", "condition"))

# Lumping all trial problems together
d$trial_status <- ifelse(d$notes == "", "Good", "Bad")

# Creating a column specifying whether a word is an adjective or noun
d$word_class <- ifelse(d$word_label %in% colours, "adjective", "noun")

# Creating a column with duration in milliseconds instead of seconds
d$dur_ms <- d$word_duration*1000

# Transforming trial to have a max value of 1
d$trial_minmax <- (d$trial - min(d$trial)) / (max(d$trial) - min(d$trial))

# Creating a data-set that is only adjectives
adj <- d[d$word_class == "adjective",]

# Removing trials with problems
adj <- adj[adj$notes == "",]
adj$effect_cat <- factor(adj$typicality,
                         levels = c("atypical", "medium", "typical"))

msa_data <- adj %>% drop_na()

# Means and SDs
saron_pictus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "saron_pictus", 1, "duration", mean(msa_data$dur_ms), sd(msa_data$dur_ms), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "saron_pictus") %>% nrow == nrow(saron_pictus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "saron_pictus") %>% pull(outcome) == saron_pictus_df %>% pull(dv)
```

### trachurus_riukiuensis

```{r}
#| label: trachurus_riukiuensis
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

#Read in output file from Praat script
msa <- read.delim(here("data", "analyses", "sc", "trachurus_riukiuensis", 
  "MSA-Praat-output.txt"), na.strings = "--undefined--")

#Read in list of NF trials (this is an edited version of one of the trial-lists on OSF)
triallist <- read.csv(here("data", "analyses", "sc", "trachurus_riukiuensis", 
  "MSA-triallist.csv"))

#Merge them together
msa2 <- merge(msa, triallist, by=c("noun","condition"))
#Adjust factor
msa2$comment <- as.factor(msa2$comment)
#Remove errors (31/900) and hesitation utterances (20/900) ~ 6% of data removed
data <- subset(msa2, comment=="NA" | comment=="noise")

#Remove JW_3; too many errors and hesitations
data <- subset(data, participant!="JW_3")

#Adjust typicality variable as dummy coded factor
data$effect_cat <- factor(data$typicality, levels = c("atypical", "medium", "typical"))

#Calculate a by participant median for each f0 measure and the median absolute deviation
#See Leys, C., Ley, C., Klein, O., Bernard, P., & Licata, L. (2013). Detecting outliers: Do not use standard deviation around the mean, use absolute deviation around the median. Journal of Experimental Social Psychology, 49(4), 764-766.
by.participant.median <- data %>%
  group_by(participant) %>%
  dplyr::summarise( n.mad = mad(noun_peakf0), n.median = median(noun_peakf0),
             adj.mad = mad(adj_peakf0), adj.median = median(adj_peakf0),
             art.mad = mad(art_peakf0), art.median = median(art_peakf0),)

#Merge by participant with full data
data.mm <- merge(by.participant.median, data)

#Create new dataframes for each f0 measure by removing less than 2.5 times median absolute deviation
noun <- subset(data.mm, noun_peakf0 > n.median-(2.5*n.mad))
#noun peak f0 removed 28 data points

adjective <- subset(data.mm, adj_peakf0 > adj.median-(2.5*adj.mad))
#adj peak f0 removed 14 data points

article <- subset(data.mm, art_peakf0 > art.median-(2*art.mad))
#art peak f0 removed 160 data points NOTE we use 2 times median absolute deviation to remove all creaky voice tokens

# Means and SDs
trachurus_riukiuensis_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "trachurus_riukiuensis", 1, "duration", mean(data$art_dur), sd(data$art_dur),
  "trachurus_riukiuensis", 2, "f0", mean(article$art_peakf0), sd(article$art_peakf0),
  "trachurus_riukiuensis", 3, "duration", mean(data$adj_dur), sd(data$adj_dur), 
  "trachurus_riukiuensis", 4, "f0", mean(adjective$adj_peakf0), sd(adjective$adj_peakf0),
  "trachurus_riukiuensis", 5, "f0", mean(noun$noun_peakf0), sd(noun$noun_peakf0),
)

# Check n models == n DVs
final_teams %>% filter(animal == "trachurus_riukiuensis") %>% nrow == nrow(trachurus_riukiuensis_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "trachurus_riukiuensis") %>% pull(outcome) == trachurus_riukiuensis_df %>% pull(dv)
```

### varanus_eulophotes

```{r}
#| label: varanus_eulophotes
#| eval: true

# Importing manually because `refit_sc.Rmd` overwrites original DVs with std DVs
# source_part(
#  file = here("scripts", "r", "04_refit", "refit_sc.Rmd"), 
#  start = , 
#  end = 
#  )

# reading in data
f0data <- read.csv(here("data", "analyses", "sc", "varanus_eulophotes", 
    "analysis", "f0data-complete.csv"), header = T)

# new dataframe with only typicality-rated data
f0data_cor <- f0data[f0data$typicality !="NA" & !is.na(f0data$typicality),]
f0data_cor$avgF0 <- as.numeric(f0data_cor$avgF0) # coercing types...
f0data_cor$effect_cat <- factor(f0data_cor$typicality, levels = c("atypical", "medium", "typical"))
msa_data <- f0data_cor

# Means and SDs
varanus_eulophotes_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "varanus_eulophotes", 1, "f0", mean(msa_data$avgF0), sd(msa_data$avgF0), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "varanus_eulophotes") %>% nrow == nrow(varanus_eulophotes_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "varanus_eulophotes") %>% pull(outcome) == varanus_eulophotes_df %>% pull(dv)
```

## TR

```{r}
#| label: tr-teams
#| echo: false
#| results: 'asis'
final_team_name_and_coder %>% 
  filter(assigned_to == "tr") %>% 
  select(animal) %>% 
  arrange(animal) %>% 
  mutate(animal = glue("- {animal}")) %>% 
  pull
```

### alosa_atun

```{r}
#| label: alosa_atun
#| eval: true
#| message: false
#| warning: false
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 353, 
  end = 507
  )

# Means and SDs
alosa_atun_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "alosa_atun", 1,  "f0",        mean(MSA_finale$meanPitch), sd(MSA_finale$meanPitch), 
  "alosa_atun", 2,  "intensity", mean(MSA_finale$maxIntensity), sd(MSA_finale$maxIntensity), 
  "alosa_atun", 3,  "duration",  mean(MSA_finale$Target.segment.duration), sd(MSA_finale$Target.segment.duration), 
  "alosa_atun", 6,  "f0",        mean(MSA_finalschwa$meanPitch), sd(MSA_finalschwa$meanPitch), 
  "alosa_atun", 7,  "intensity", mean(MSA_finalschwa$maxIntensity), sd(MSA_finalschwa$maxIntensity), 
  "alosa_atun", 8,  "duration",  mean(MSA_finalschwa$Target.segment.duration), sd(MSA_finalschwa$Target.segment.duration), 
  "alosa_atun", 11, "duration",  mean(MSA_both$DiphDur), sd(MSA_both$DiphDur), 
  "alosa_atun", 12, "duration",  mean(MSA_both$DurRat), sd(MSA_both$DurRat), 
  "alosa_atun", 14, "duration",  mean(MSA_finale$Target.word.duration), sd(MSA_finale$Target.word.duration), 
  "alosa_atun", 15, "duration",  mean(MSA_finale$Pause.duration, na.rm = T), sd(MSA_finale$Pause.duration, na.rm = T), 
  "alosa_atun", 16, "duration",  mean(MSA_both$syllablesPerMinute), sd(MSA_both$syllablesPerMinute), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "alosa_atun") %>% nrow == nrow(alosa_atun_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "alosa_atun") %>% pull(outcome) == alosa_atun_df %>% pull(dv)
```

### chelonia_brummeri

```{r}
#| label: chelonia_brummeri
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 1532, 
  end = 1601
  )

# Means and SDs
chelonia_brummeri_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "chelonia_brummeri", 1,  "duration",  mean(d$rel_dur), sd(d$rel_dur), 
  "chelonia_brummeri", 2,  "intensity", mean(d$rel_intens), sd(d$rel_intens), 
  "chelonia_brummeri", 3,  "f0",        mean(d$f0range_st), sd(d$f0range_st), 
  "chelonia_brummeri", 4,  "f0",        mean(d$adj_maxf0), sd(d$adj_maxf0), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "chelonia_brummeri") %>% nrow == nrow(chelonia_brummeri_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "chelonia_brummeri") %>% pull(outcome) == chelonia_brummeri_df %>% pull(dv)
```

### dunkleosteus_inscriptus

```{r}
#| label: dunkleosteus_inscriptus
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 1309, 
  end = 1420
  )

# Means and SDs
dunkleosteus_inscriptus_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "dunkleosteus_inscriptus", 1,  "duration", mean(dat$rel_n), sd(dat$rel_n), 
  "dunkleosteus_inscriptus", 2,  "duration", mean(dat$dur_n), sd(dat$dur_n), 
  "dunkleosteus_inscriptus", 3,  "duration", mean(dat$rel_a), sd(dat$rel_a), 
  "dunkleosteus_inscriptus", 4,  "duration", mean(dat$dur_a), sd(dat$dur_a), 
  "dunkleosteus_inscriptus", 5,  "duration", mean(dat$dur_np), sd(dat$dur_np),
)

# Check n models == n DVs
final_teams %>% filter(animal == "dunkleosteus_inscriptus") %>% nrow == nrow(dunkleosteus_inscriptus_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "dunkleosteus_inscriptus") %>% pull(outcome) == dunkleosteus_inscriptus_df %>% pull(dv)
```

### pervagor_meeki

```{r}
#| label: pervagor_meeki
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 896, 
  end = 949
  )

# Means and SDs
pervagor_meeki_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "pervagor_meeki", 1,  "intensity", mean(data_int_trim[data_int_trim$word_type == "target_name",]$Overall_Intensity), sd(data_int_trim[data_int_trim$word_type == "target_name",]$Overall_Intensity), 
  "pervagor_meeki", 2,  "duration", mean(data_dur_trim[data_dur_trim$word_type == "target_name",]$Duration_ms), sd(data_dur_trim[data_dur_trim$word_type == "target_name",]$Duration_ms), 
  "pervagor_meeki", 3,  "f0", mean(data_meanf0_trim[data_meanf0_trim$word_type == "target_name",]$Mean_f0_st), sd(data_meanf0_trim[data_meanf0_trim$word_type == "target_name",]$Mean_f0_st), 
  "pervagor_meeki", 4,  "f0", mean(data_maxf0_trim[data_maxf0_trim$word_type == "target_name",]$Max_f0_st), sd(data_maxf0_trim[data_maxf0_trim$word_type == "target_name",]$Max_f0_st), 
  "pervagor_meeki", 5,  "intensity", mean(data_int_trim[data_int_trim$word_type == "target_colour",]$Overall_Intensity), sd(data_int_trim[data_int_trim$word_type == "target_colour",]$Overall_Intensity), 
  "pervagor_meeki", 6,  "duration",  mean(data_dur_trim[data_dur_trim$word_type == "target_colour",]$Duration_ms), sd(data_dur_trim[data_dur_trim$word_type == "target_colour",]$Duration_ms),
  "pervagor_meeki", 7,  "f0", mean(data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",]$Mean_f0_st), sd(data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",]$Mean_f0_st),
  "pervagor_meeki", 8,  "f0", mean(data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",]$Max_f0_st), sd(data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",]$Max_f0_st),
)

# Check n models == n DVs
final_teams %>% filter(animal == "pervagor_meeki") %>% nrow == nrow(pervagor_meeki_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "pervagor_meeki") %>% pull(outcome) == pervagor_meeki_df %>% pull(dv)
```

### procambarus_mahogoni

```{r}
#| label: procambarus_mahogoni
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 1742, 
  end = 1745
  )

# Means and SDs
procambarus_mahogoni_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "procambarus_mahogoni", 1,  "duration",  mean(msa_data_raw$v1_duration_c), sd(msa_data_raw$v1_duration_c), 
  "procambarus_mahogoni", 2,  "f0",        mean(msa_data_raw$v1_pitch_mean_c), sd(msa_data_raw$v1_pitch_mean_c), 
  "procambarus_mahogoni", 3,  "intensity", mean(msa_data_raw$v1_int_mean_c), sd(msa_data_raw$v1_int_mean_c), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "procambarus_mahogoni") %>% nrow == nrow(procambarus_mahogoni_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "procambarus_mahogoni") %>% pull(outcome) == procambarus_mahogoni_df %>% pull(dv)
```

### trachyphyllia_lappa

```{r}
#| label: trachyphyllia_lappa
#| eval: true
source_part(
  file = here("scripts", "r", "04_refit", "refit_tr.Rmd"), 
  start = 49, 
  end = 134
  )

duration_adj_data <- durations %>% filter(word == 9) %>% mutate(type = "adjective")
duration_noun_data <- durations %>% filter(word == 10) %>% mutate(type = "noun")
pitch_adj_data <- pitchfile %>% filter(word == 9) %>% mutate(type = "adjective")
pitch_noun_data <- pitchfile %>% filter(word == 10) %>% mutate(type = "noun")

# Means and SDs
trachyphyllia_lappa_df <- tribble(
  ~"animal", ~"model_n", ~"dv", ~"mean", ~"sd", 
  "trachyphyllia_lappa", 1,  "duration", mean(duration_adj_data$duration), sd(duration_adj_data$duration), 
  "trachyphyllia_lappa", 2,  "f0",       mean(pitch_adj_data$maximum), sd(pitch_adj_data$maximum), 
  "trachyphyllia_lappa", 3,  "f0", mean(pitch_adj_data$minimum), sd(pitch_adj_data$minimum), 
  "trachyphyllia_lappa", 4,  "f0", mean(pitch_adj_data$range), sd(pitch_adj_data$range), 
  "trachyphyllia_lappa", 5,  "duration", mean(duration_noun_data$duration), sd(duration_noun_data$duration), 
  "trachyphyllia_lappa", 6,  "f0", mean(pitch_noun_data$maximum), sd(pitch_noun_data$maximum), 
  "trachyphyllia_lappa", 7,  "f0", mean(pitch_noun_data$minimum), sd(pitch_noun_data$minimum), 
  "trachyphyllia_lappa", 8,  "f0", mean(pitch_noun_data$range), sd(pitch_noun_data$range), 
)

# Check n models == n DVs
final_teams %>% filter(animal == "trachyphyllia_lappa") %>% nrow == nrow(trachyphyllia_lappa_df)

# Check DVs are the same as those listen in model excel
final_teams %>% filter(animal == "trachyphyllia_lappa") %>% pull(outcome) == trachyphyllia_lappa_df %>% pull(dv)
```

# Build table

```{r}
#| label: build-tibble

# Comnine tibbles 
msa_mean_sd <- bind_rows(
  anomalocaris_ornata_df, 
  anthracoceros_coronata_df, 
  arapaima_modularis_df, 
  clione_dorsalis_df, 
  comanthina_maculatus_df, 
  cromileptes_saxatilis_df, 
  gnathosaurus_canadensis_df, 
  gymnothorax_spinulosus_df, 
  haematopus_fossor_df, 
  lasionycteris_altavela_df, 
  naso_cassivellaunos_df, 
  psittacula_scabriculus_df, 
  trapezia_cantonensis_df, 
  dermatolepis_aculeatus_df, 
  eriphia_laterispinis_df, 
  hoplostethus_macrosteus_df, 
  neosilurus_omanensis_df, 
  nestor_idahoensis_df, 
  paralichthys_undulatus_df, 
  petauroides_fistulator_df, 
  saron_pictus_df, 
  trachurus_riukiuensis_df, 
  varanus_eulophotes_df, 
  alosa_atun_df, 
  chelonia_brummeri_df, 
  dunkleosteus_inscriptus_df, 
  pervagor_meeki_df, 
  procambarus_mahogoni_df, 
  trachyphyllia_lappa_df
)

msa_mean_sd_df <- left_join(
  msa_mean_sd %>% 
    arrange(animal) %>% 
    mutate(row_n = row_number()), 
  assigned %>%
    filter(
      animal %in% teams, 
      exclude == "no", 
      outcome %in% c("duration", "intensity", "f0")
    ) %>% 
    select(animal, model_n, outcome, transformation) %>% 
    arrange(animal) %>% 
    mutate(row_n = row_number()), 
  by = c("animal", "row_n")
) %>% 
  mutate(model_n_match = if_else(model_n.x == model_n.y, 1, 0), 
    transformation = case_when(
      is.na(transformation) ~ 'none', 
      transformation == "NA" ~ 'none', 
      TRUE ~ transformation
    )) %>% 
  select(animal, model_n_rda = model_n.x, model_n_excel = model_n.y, model_n_match, 
         dv, outcome, transformation, mean, sd)

# Save as .csv
if(F) {
write_csv(msa_mean_sd_df, here("data", "analyses", "msa_mean_sd_df.csv"))
}

# Print to html for funsies
msa_mean_sd_df %>% 
  knitr::kable()
```