### 2022-09-10 scripts/r/03_analyses_coding.Rmd
♻️ - split 04_analyses_coding, update msa_models

---
title: "03 - Analyses coding"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
#library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(ggmosaic)
library(googlesheets4)
library(magrittr)
library(glue)
library(osfr)
library(cli)
library("here")
library("fs")
library("forcats")
library("strex")
library("purrr")
library("broom")
library("broom.mixed")

# Plotting theme
clean_theme <- function(...) {
  list(
    theme_minimal(), 
    theme(
      axis.title.x = element_blank(),
      panel.grid.major = element_line(colour = 'grey90', size = 0.15),
      panel.grid.minor = element_line(colour = 'grey90', size = 0.15),
      strip.text.x = element_text(hjust = 0),
      axis.text.x = element_blank()
    )
  )
}
```

# Read coding sheet

```{r coding}
the_sheet <- "https://docs.google.com/spreadsheets/d/1OOXB-8Uk-fh_urw0Lm4DC0jBueXLIb4Bm-a9--UVXOw/edit?usp=sharing"

coding <- read_sheet(the_sheet) %>%
  drop_na(model_n)
```

# Read intake forms

We will join the intake forms with `msa_models`.

```{r intake}
other_sheet <- "https://docs.google.com/spreadsheets/d/1KAZC2vyteTfGE6mlwOiz9b8F9VghDYLGsOzmWaNbmEE/edit?usp=sharing"

team_animal_coord <- read_sheet(other_sheet)

msa_intake <- read_csv("./data/questionnaires/msa_intake_form.csv") %>%
  filter(row_number() != 2)
msa_intake_qs <- msa_intake[1,]
msa_intake <- msa_intake[-1,]

msa_intake <- msa_intake %>%
  mutate(
    Q6 = case_when(
      Q6 == "Zlatomira IIchovska" ~ "Zlatomira Ilchovska",
      Q6 == "University of Birmingham team under the team coordinator Zlatomira IIchovska" ~ "Zlatomira Ilchovska",
      Q6 == "Vincent Hughes" ~ "Vince Hughes",
      Q6 == "Tiphaine Cordelier" ~ "Tiphaine Caudrelier",
      Q6 == "Scott Perry" ~ "Scott James Perry",
      Q6 == "Scott James Perru" ~ "Scott James Perry",
      Q6 == "Nicole Rodriquez" ~ "Nicole Rodriguez",
      Q6 == "Michael Gradoville (Arizona State Univ.)" ~ "Michael Gradoville",
      Q6 == "Joey Stanley" ~ "Joseph A. Stanley",
      Q6 == "Dusan Nikolic (dusan.nikolic1@ucalgary.ca)" ~ "Dušan Nikolić",
      Q6 == "Dusan Nikolic" ~ "Dušan Nikolić",
      Q6 == "Dušan Nickolić" ~ "Dušan Nikolić",
      Q6 == "Alicea Megan Brown" ~ "Alicia M Brown",
      Q6 == "Alicia Brown" ~ "Alicia M Brown",
      Q6 == "Ali A-Hoorie" ~ "Ali H. Al-Hoorie",
      Q6 == "Ali Al-Hoorie" ~ "Ali H. Al-Hoorie",
      Q6 == "Dr Rory Turnbull" ~ "Rory Turnbull",
      Q6 == "Dr. Reenu Punnoose" ~ "Reenu Punnoose",
      TRUE ~ Q6
    ),
    Q5 = as.numeric(Q5),
    Q7_1 = as.numeric(Q7_1)
  )

msa_intake_teams <- msa_intake %>%
  group_by(Q6) %>%
  summarise(
    years_from_phd = mean(Q5, na.rm = TRUE),
    years_from_phd_sd = sd(Q5, na.rm = TRUE),
    prior_belief = mean(Q7_1, na.rm = TRUE),
    prior_belief_sd = sd(Q7_1, na.rm = TRUE)
  ) %>%
  full_join(y = team_animal_coord, by = c("Q6" = "coordinator")) %>%
  select(-Q6)
```

# Read reviews

```{r}
msa_reviews <- readRDS("./data/peer_review/msa_reviews.rds")

msa_reviews_teams <- msa_reviews %>%
  group_by(Q3) %>%
  summarise(
    phon_rating = mean(Q5_2, na.rm = TRUE),
    phon_sd = sd(Q5_2, na.rm = TRUE),
    stat_rating = mean(Q5_3, na.rm = TRUE),
    stat_sd = sd(Q5_3, na.rm = TRUE),
    all_rating = mean(Q5_4, na.rm = TRUE),
    all_sd = sd(Q5_4, na.rm = TRUE),
  )
```

# Teams' analyses

## Download OSF components

First, we get a list of the OSF components in the Teams Analyses component. (It takes a few seconds)

```{r comps}
assigned <- coding %>%
  dplyr::select(team, animal, assigned_to) %>%
  distinct()

comps <- osf_retrieve_node("https://osf.io/n3fyd/") %>%
  osf_ls_nodes(n_max = Inf) %>%
  left_join(y = assigned, by = c("name" = "animal"))
```

Now we get only those components that are assigned to you. Tell me who you are when prompted.

```{r who}
who <- menu(c("Ste", "Joseph", "Timo"), "Who are you?")

whos <- c("sc", "jvc", "tr")
your_comps <- comps %>%
  filter(assigned_to == whos[who])
dir.create(glue("./data/analyses/{whos[who]}"), showWarnings = FALSE, recursive = TRUE)
```

And now we download the components in `data/analyses/` (it will take some time!).

```{r download}
for (comp in 1:nrow(your_comps)) {
  name <- your_comps$name[comp]
  files <- osf_ls_files(your_comps[comp,], n_max = Inf)
  dir <- glue("./data/analyses/{whos[who]}/{name}")
  dir.create(dir, showWarnings = FALSE, recursive = TRUE)
  cli_h1(name)
  osf_download(
    files,
    path = dir,
    recurse = TRUE,
    conflicts = "skip",
    verbose = TRUE, progress = TRUE
  )
}
```

## Refit models

Now that the components have been downloaded, we can refit the models using the scripts in `scripts/r/04-refit/`.