### 2022-10-17 scripts/r/08_draft_comments.Rmd

---
title: "Get co-author comments on draft"
author: "Joseph V. Casillas"
output: html_document
date: "Last upate: `r Sys.Date()`"
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE
  )

knitr::opts_knit$set(root.dir = here::here())
```

```{r}
#| label: load-libs
library("googledrive")
library("xml2")
library("dplyr")
library("tidyr")
library("ggplot2")
library("forcats")
library("readr")
library("kableExtra")
```

```{r}
#| label: identify-download
# give gargle permission
drive_auth(email = TRUE)

# Locate file, check ID
drive_find(n_max = 5, type = "document")

# Search by id (just checking)
drive_get(id = "1CFgRo93mRgifpuFOuQE3vNBeMW-H7ps9eD--vxH-6CQ")

# Download to doc folder
drive_download(
  file = "Draft_RR", 
  path = "./RR_manuscript/commented_draft/Draft_RR.docx", 
  type = "docx",
  overwrite = T
  )
```

```{r}
#| label: unzip-word
# Get word doc and unzip to 'unzipped' folder, will create if doesn't exist
system("unzip -o ./RR_manuscript/commented_draft/Draft_RR.docx -d ./RR_manuscript/commented_draft/unzipped/")
```

```{r}
#| label: get-xml
# Load xml file
comments_xml <- read_xml("./RR_manuscript/commented_draft/unzipped/word/comments.xml")

# Explore structure
xml_name(comments_xml)
xml_children(comments_xml) %>% head
# xml_structure(comments_xml) 
# Key structure is: <comment [author, id, date]>

# Get all comments
comments <- xml_find_all(comments_xml, ".//w:comment")
```

```{r}
#| label: create-df
# Create df
comments_df <- tibble(
  commenter = trimws(xml_attr(comments, "author")), 
  date = trimws(xml_attr(comments, "date")), 
  comment = trimws(xml_text(comments)), 
  type = "", 
  assigned = "", 
  solution = "", 
  done = "", 
  notes = ""
  ) %>% 
  arrange(commenter, date) %>% 
  write_csv("./data/rr_comments/comment_log.csv")
```

```{r}
#| label: comment-descriptives

# N commenters
comments_df$commenter %>% unique() %>% length

# N comments
nrow(comments_df)

# N comments per commenter
comments_df %>% 
  count(commenter) %>% 
  mutate(commenter = fct_reorder(commenter, n, min)) %>% 
  ggplot() + 
  aes(x = n, y = commenter) + 
  geom_segment(aes(x = 0, xend = n, yend = commenter)) + 
  geom_point() 
```

```{r}
#| label: print-df
comments_df %>% 
  kbl() %>%
  kable_paper("hover", full_width = F) %>% 
  scroll_box(height = "500px")
```

```{r}
#| label: upload-to-drive
#| eval: false
drive_upload(
  media = "./data/rr_comments/comment_log.csv", 
  path = "many_analyses_docs/rr_submission", 
  name = "comment_log", 
  type = "spreadsheet", 
  overwrite = TRUE
)
```