### 2023-01-16 scripts/r/04_refit/refit_tr.Rmd
🐛 - include pseudopleuronectes_assasi

---
title: "Re-fit"
subtitle: "TR"
output: html_document
date: "Last update: `r Sys.Date()`"
---

```{r setup}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
knitr::opts_knit$set(root.dir = here::here())

# libs
library(tidyverse)
library(brms)
library(readxl)
library(lme4)
library(lmerTest)
library(here)

# NOTE: one of the many packages interferes with tidyverse, so the select function throws an error!

# Plotting theme because Joseph is extra
clean_theme <- function(...) {
  list(
    theme_minimal(), 
    theme(
      axis.title.y = element_text(size = rel(.9), hjust = 0.95),
      axis.title.x = element_text(size = rel(.9), hjust = 0.95),
      panel.grid.major = element_line(colour = 'grey90', size = 0.15),
      panel.grid.minor = element_line(colour = 'grey90', size = 0.15))
  )
}

# Functional sequence to get relevant info from Rds files
clean_up <- . %>% 
  tidy(effects = "fixed") %>% 
  suppressWarnings() %>% 
  filter(term %in% c("effect_cattypical", "effect_con")) %>% 
  dplyr::select(term, estimate, se = std.error)

```

# trachyphyllia_lappa

```{r trachyphyllia_lappa_setup}

path <- here("data", "analyses", "tr", "trachyphyllia_lappa")

path_trial = paste0(path, "/trial_level_data/")
msa <- list.files(path = path_trial, pattern = "*.csv", full.names = TRUE) %>% 
  lapply(read_csv,  col_types = cols( .default = col_character())) %>% 
  bind_rows %>%
  mutate(trial = as.numeric(trial))

# analyze only NF condition with the first 35 trials
msa_nf = msa %>% filter(condition == "NF" & trial < 36)

results_path = paste0(path, "/acoustic_measurements/errors.txt")
errorfile = read.delim(results_path, skipNul = TRUE, fileEncoding="latin1", header = TRUE) %>%
  mutate(label = as.factor(label)) %>%
  group_by(speaker) %>%
  mutate(trial = as.integer(row_number()/2))

main_errors = errorfile %>% filter(!label %in% "") %>% filter(trial < 36) %>%
  filter(!label %in% c("team_add", "team_delete") )

main_errors %>% group_by(label) %>% summarise(trials = n()) %>%
       mutate(percent = trials/sum(trials))

results_path = paste0(path, "/acoustic_measurements/durations.txt")
durations = read.delim(results_path,skipNul = TRUE, fileEncoding="latin1", header = TRUE) 
newnames = c("speaker", "phrase", "duration")

durations = durations %>%
  rename_with(~ newnames[which(names(durations) == .x)], .cols = names(durations))%>%
  group_by(speaker) %>%
  mutate(trial = as.integer(gl(n(), 11, n()))) %>% 
  group_by(speaker, trial) %>%
  mutate(word = row_number()) %>%
filter(word %in% c(9,10)) %>%
  ## merge with msa data
  left_join(msa_nf %>% select(speaker,trial, condition, typicality, target_colour) %>%
              mutate(trial = as.integer(trial))) %>%
  filter(!is.na(condition))

# combine with phrase info
phrase_data = durations %>% select(speaker, trial, phrase) %>%
  group_by(speaker, trial) %>%
   summarise(phrase = paste(phrase, collapse = "-"))

durations = durations %>% select(-phrase) %>%
  left_join(phrase_data)

# exclude error trials
durations = durations %>% left_join(main_errors) %>%
  filter(is.na(label))

durations %>% group_by(speaker) %>% summarise(n=n()/2) %>%
  summarise(mean_trials = mean(n), sd_trials = sd(n))

# pitch extraction
results_path = paste0(path,"/acoustic_measurements/pitchresults.txt")
pitchfile = read.delim(results_path, skipNul = TRUE, fileEncoding="latin1", header = TRUE)
newnames = c("speaker", "phrase", "maximum","minimum")

pitchfile = pitchfile %>%
  rename_with(~ newnames[which(names(pitchfile) == .x)], .cols = names(pitchfile))%>%
  group_by(speaker) %>%
  mutate(trial = as.integer(gl(n(), 11, n()))) %>% 
  group_by(speaker, trial) %>%
  mutate(word = row_number()) %>%
filter(word %in% c(9,10)) %>%
  ## merge with msa data
  left_join(msa_nf %>% select(speaker,trial, condition, typicality, target_colour) %>%
              mutate(trial = as.integer(trial))) %>%
  filter(!is.na(condition)) %>%
  pivot_longer(names_to = "pitch", cols = minimum:maximum) %>%
  mutate(value = ifelse(value == "--undefined--", NA, value),
         value = as.numeric(value))

pitchfile = pitchfile %>% select(-phrase) %>%
  left_join(phrase_data)

# exclude errors
pitchfile = pitchfile %>% left_join(main_errors) %>%
  filter(is.na(label)) %>%
# pitch into wide 
  pivot_wider(names_from = pitch, values_from = value)%>%
  mutate(range = maximum - minimum)

```

```{r trachyphyllia_lappa_models}

ctrl = lmerControl(check.conv.singular = .makeCC(action = "ignore",  tol = 1e-4),
                   optimizer = "optimx", optCtrl  = list(method="bobyqa", maxfun = 2e+6))

duration_int = duration ~ typicality*target_colour + 
  (typicality + target_colour  | speaker) + (1 | phrase)
duration_noint = duration ~ typicality + target_colour + 
  (typicality + target_colour  | speaker) + (1 | phrase)

pitch_min_int = minimum ~ typicality * target_colour + (typicality + target_colour  | speaker)+ 
  (1 | phrase)
pitch_min_noint = minimum ~ typicality + target_colour + (typicality + target_colour  | speaker)+ 
  (1 | phrase)

pitch_max_int = maximum ~ typicality * target_colour + (typicality + target_colour  | speaker)+
   (1 | phrase)
pitch_max_noint = maximum ~ typicality + target_colour + (typicality + target_colour  | speaker)+
   (1 | phrase)

pitch_range_noint = range ~ typicality + target_colour + (typicality + target_colour  | speaker)+
   (1 | phrase)
pitch_range_int = range ~ typicality * target_colour + (typicality + target_colour  | speaker)+
   (1 | phrase)

# adjective models - duration
duration_adj_data = durations %>% filter(word == 9) %>% 
  mutate(type = "adjective")

# TR: no interaction better fit, so chosen by the analysts
duration_adj_model_noint = lme4::lmer(data = duration_adj_data, duration_noint, control=ctrl)

# nouns models - duration
duration_noun_data = durations %>% filter(word == 10) %>% mutate(type = "noun")

# TR: interaction model picked
duration_noun_model_int =lme4::lmer(data = duration_noun_data, duration_int, control=ctrl)

# adjective model - pitch
pitch_adj_data = pitchfile %>% filter(word == 9) %>% mutate(type = "adjective")
pitch_noun_data = pitchfile %>% filter(word == 10) %>% mutate(type = "noun")

pitch_adj_model_max_noint = lme4::lmer(data = pitch_adj_data,pitch_max_noint, control = ctrl) 
pitch_adj_model_min_noint = lme4::lmer(data = pitch_adj_data, pitch_min_noint, control = ctrl)
pitch_adj_model_range_noint = lme4::lmer(data = pitch_adj_data, pitch_range_noint, control = ctrl)

pitch_noun_model_max_noint = lme4::lmer(data = pitch_noun_data,pitch_max_noint, control = ctrl) 
pitch_noun_model_min_noint = lme4::lmer(data = pitch_noun_data, pitch_min_noint, control = ctrl)
pitch_noun_model_range_noint = lme4::lmer(data = pitch_noun_data, pitch_range_noint, control = ctrl)

```

```{r trachyphyllia_lappa_fitting}

# standardize
duration_adj_data <- duration_adj_data %>% 
  mutate(
    #duration_s = (duration - mean(duration, na.rm = TRUE)) / sd(duration,  na.rm = TRUE),
    # does not work and I don't know why. 
    # dummy-code typicality
    typicality = as.factor(typicality),
    target_colour_s = as.factor(target_colour),
    effect_cat = C(typicality, treatment))

contrasts(duration_adj_data$target_colour_s) <- contr.sum(5)

duration_adj_data$duration_s = (duration_adj_data$duration - mean(duration_adj_data$duration, na.rm = TRUE)) / sd(duration_adj_data$duration,  na.rm = TRUE)

duration_noun_data <- duration_noun_data %>% 
  mutate(
         # scale continuous variables
         # duration_s = (duration - mean(duration, na.rm = TRUE)) / sd(duration,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment)
  )

contrasts(duration_noun_data$target_colour_s) <- contr.sum(5)

duration_noun_data$duration_s = (duration_noun_data$duration - mean(duration_noun_data$duration, na.rm = TRUE)) / sd(duration_noun_data$duration,  na.rm = TRUE)

pitch_adj_data <- pitch_adj_data %>% 
  mutate(
         # scale continuous variables
         # minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         # maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         # range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment)
  )

contrasts(pitch_adj_data$target_colour_s) <- contr.sum(5)

pitch_adj_data$minimum_s = (pitch_adj_data$minimum - mean(pitch_adj_data$minimum, na.rm = TRUE)) / sd(pitch_adj_data$minimum,  na.rm = TRUE)
pitch_adj_data$maximum_s = (pitch_adj_data$maximum - mean(pitch_adj_data$maximum, na.rm = TRUE)) / sd(pitch_adj_data$maximum,  na.rm = TRUE)
pitch_adj_data$range_s = (pitch_adj_data$range - mean(pitch_adj_data$range, na.rm = TRUE)) / sd(pitch_adj_data$range,  na.rm = TRUE)

pitch_noun_data <- pitch_noun_data %>% 
  mutate(
         # scale continuous variables
         #minimum_s = (minimum - mean(minimum, na.rm = TRUE)) / sd(minimum,  na.rm = TRUE),
         #maximum_s = (maximum - mean(maximum, na.rm = TRUE)) / sd(maximum,  na.rm = TRUE),
         #range_s = (range - mean(range, na.rm = TRUE)) / sd(range,  na.rm = TRUE),
         # dummy-code typicality
         typicality = as.factor(typicality),
         target_colour_s = as.factor(target_colour),
         effect_cat = C(typicality, treatment),
         # sum-code categorical predictors
         # target_colour_s = C(target_colour, sum, 5)
  )

# sum-code categorical predictors
contrasts(pitch_noun_data$target_colour_s) <- contr.sum(5)

pitch_noun_data$minimum_s = (pitch_noun_data$minimum - mean(pitch_noun_data$minimum, na.rm = TRUE)) / sd(pitch_noun_data$minimum,  na.rm = TRUE)
pitch_noun_data$maximum_s = (pitch_noun_data$maximum - mean(pitch_noun_data$maximum, na.rm = TRUE)) / sd(pitch_noun_data$maximum,  na.rm = TRUE)
pitch_noun_data$range_s = (pitch_noun_data$range - mean(pitch_noun_data$range, na.rm = TRUE)) / sd(pitch_noun_data$range,  na.rm = TRUE)

# run brms equivalents
trachyphyllia_lappa_1_duradj_cat <- brm(
  duration_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = duration_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_1_duradj_cat"
  )

trachyphyllia_lappa_2_f0maxadj_cat <- brm(
  maximum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_2_f0maxadj_cat"
  )

trachyphyllia_lappa_3_f0minadj_cat <- brm(
  minimum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) +  
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_3_f0minadj_cat"
  )

trachyphyllia_lappa_4_f0rangeadj_cat <- brm(
  range_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_adj_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_4_f0rangeadj_cat"
  )

# refitted with more samples because of the interaction
trachyphyllia_lappa_5_durnoun_cat <- brm(
  duration_s ~  effect_cat * target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = duration_noun_data, 
  iter = 4000,
  control = list(adapt_delta = 0.9999),
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_5_durnoun_cat"
  )

trachyphyllia_lappa_6_f0maxnoun_cat <- brm(
  maximum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_6_f0maxnoun_cat"
  )

trachyphyllia_lappa_7_f0minnoun_cat <- brm(
  minimum_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) +  
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_7_f0minnoun_cat"
  )

trachyphyllia_lappa_8_f0rangenoun_cat <- brm(
  range_s ~  effect_cat + target_colour_s + 
      (typicality + target_colour_s  | speaker) + 
      (1 | phrase),
  data = pitch_noun_data, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/trachyphyllia_lappa_8_f0rangenoun_cat"
  )

```

# alosa_atun

```{r alosa_atun_setup}

MSA_combined <- readRDS(here("data", "analyses", "tr", "alosa_atun", "alosa_atun_data.rds"), refhook = NULL)

MSA_combined$meanPitch = as.numeric(MSA_combined$meanPitch)

MSA_combined_all <-
MSA_combined %>%
  filter(Text == "der",
         After.Match != "Würfel") %>%
  filter(!is.na(meanPitch))

MSA_combined_all <-
MSA_combined_all %>%
  filter(!is.na(Target.word.duration))

MSA_combined_all <-
MSA_combined_all %>%
  filter(!is.na(condition)) %>%
  filter(condition != "NA")

# create summary of F1 and F2 values based on speaker and vowels. sd_limit, max and min variables are used for outlier removal in the next step

sd_limit = 2.5

#set this to the +/- sd limit you want to filter outliers by

MSA_summstats <- MSA_combined_all %>%
  group_by(Transcript, Target.segment) %>%
  summarise(mean_F1 = mean(f1_time_0_5),
            mean_F2 = mean(f2_time_0_5),
            mean_amp = mean(maxIntensity),
            mean_pitch = mean(meanPitch),
            mean_durSeg = mean(Target.segment.duration),
            mean_durWord = mean(Target.word.duration),
            sd_F1 = sd(f1_time_0_5),
            sd_F2 = sd(f2_time_0_5),
            sd_amp = sd(maxIntensity),
            sd_pitch = sd(meanPitch),
            sd_durSeg = sd(Target.segment.duration),
            sd_durWord = sd(Target.word.duration),
            max_F1 = mean(f1_time_0_5) + sd_limit*(sd(f1_time_0_5)),
            min_F1 = mean(f1_time_0_5) - sd_limit*(sd(f1_time_0_5)),
            max_F2 = mean(f2_time_0_5) + sd_limit*(sd(f2_time_0_5)),
            min_F2 = mean(f2_time_0_5) - sd_limit*(sd(f2_time_0_5)),
            max_amp = mean(maxIntensity) + sd_limit*(sd(maxIntensity)),
            min_amp = mean(maxIntensity) - sd_limit*(sd(maxIntensity)),
            max_pitch = mean(meanPitch) + sd_limit*(sd(meanPitch)),
            min_pitch = mean(meanPitch) - sd_limit*(sd(meanPitch)),
            max_durSeg = mean(Target.segment.duration) + sd_limit*(sd(Target.segment.duration)),
            min_durSeg = mean(Target.segment.duration) - sd_limit*(sd(Target.segment.duration)),
            max_durWord = mean(Target.word.duration) + sd_limit*(sd(Target.word.duration)),
            min_durWord = mean(Target.word.duration) - sd_limit*(sd(Target.word.duration)))

#store the outlier tokens data

outlier_tokens <- MSA_combined_all %>%
  left_join(., MSA_summstats) %>%
  mutate(outlier = ifelse(f1_time_0_5 > min_F1 &
                            f1_time_0_5 < max_F1 &
                            f2_time_0_5 > min_F2 &
                            f2_time_0_5 < max_F2 &
                            maxIntensity > min_amp &
                            maxIntensity < max_amp &
                            meanPitch > min_pitch &
                            meanPitch < max_pitch &
                            Target.segment.duration > min_durSeg &
                            Target.segment.duration < max_durSeg &
                            Target.word.duration > min_durWord &
                            Target.word.duration < max_durWord &
                            f2_time_0_5 > 800,  
                          FALSE,
                          TRUE)) %>%
  group_by(Transcript, Target.segment) %>%
  filter(outlier == TRUE) %>%
  ungroup()

#add the summary statistics and filter out outliers

MSA_combined_pruned <- MSA_combined_all %>%
  left_join(., MSA_summstats) %>%
  mutate(outlier = ifelse(f1_time_0_5 > min_F1 &
                            f1_time_0_5 < max_F1 &
                            f2_time_0_5 > min_F2 &
                            f2_time_0_5 < max_F2 &
                            maxIntensity > min_amp &
                            maxIntensity < max_amp &
                            meanPitch > min_pitch &
                            meanPitch < max_pitch &
                            Target.segment.duration > min_durSeg &
                            Target.segment.duration < max_durSeg &
                            Target.word.duration > min_durWord &
                            Target.word.duration < max_durWord &
                            f2_time_0_5 > 800,  
                          FALSE,
                          TRUE)) %>%
  group_by(Transcript, Target.segment) %>%
  filter(outlier == FALSE) %>%
  ungroup()

#get new summary values
MSA_summstats <- MSA_combined_pruned %>%
  group_by(Transcript, Target.segment) %>%
  summarise(n_tokens_vowel = n(),
            mean_F1 = mean(f1_time_0_5),
            mean_F2 = mean(f2_time_0_5),
            mean_amp = mean(maxIntensity),
            mean_durSeg = mean(Target.segment.duration),
            mean_durWord = mean(Target.word.duration),
            mean_pitch = mean(meanPitch),
            sd_F1 = sd(f1_time_0_5),
            sd_F2 = sd(f2_time_0_5),
            sd_amp = sd(maxIntensity),
            sd_durSeg = sd(Target.segment.duration),
            sd_durWord = sd(Target.word.duration),
            sd_pitch = sd(meanPitch)) %>%
  ungroup()

MSA_controls <-
MSA_combined %>%
  filter(Text == "den") %>%
  filter(Target.segment == "e" | Target.segment == "@" ) %>%
  filter(After.Match == "Würfel")

MSA_combined_pruned$item = paste(MSA_combined_pruned$Transcript, MSA_combined_pruned$trialnumber, MSA_combined_pruned$trialtype)

MSA_controls$item = paste(MSA_controls$Transcript, MSA_controls$trialnumber, MSA_controls$trialtype)

MSA_controls_selected = MSA_controls %>% select(item, meanPitch, maxIntensity)
names(MSA_controls_selected) = c("item", "controlPitch", "controlIntensity")

MSA_final = left_join(MSA_combined_pruned, MSA_controls_selected, by="item")

MSA_final$Pause.duration = MSA_final$Token.plus.1.word.start - MSA_final$Target.word.end

MSA_finale <-
  MSA_final %>%
  filter(Target.segment == "e")

MSA_finalschwa <-
  MSA_final %>%
  filter(Target.segment == "@")

MSA_finale$eduration = MSA_finale$Target.segment.duration
MSA_finalschwa$schwaduration = MSA_finalschwa$Target.segment.duration
MSA_finalschwa$schwaF1 = MSA_finalschwa$f1_time_0_5
MSA_finalschwa$schwaF2 = MSA_finalschwa$f2_time_0_5

MSA_interimschwa = MSA_finalschwa %>% select(item, schwaduration, schwaF1, schwaF2)
MSA_both = left_join(MSA_finale, MSA_interimschwa, by="item")
MSA_both = MSA_both %>% filter(!is.na(schwaduration))
MSA_both$DiphDur = MSA_both$Target.segment.duration + MSA_both$schwaduration
MSA_both$DurRat = MSA_both$Target.segment.duration / MSA_both$schwaduration
MSA_both$ED = sqrt((MSA_both$f1_time_0_5-MSA_both$schwaF1)^2 + (MSA_both$f2_time_0_5-MSA_both$schwaF2)^2)

```

```{r alosa_atun_models}

# schwa analyses
model_f1c <-
MSA_finale %>%
  lmer(., formula=f1_time_0_5 ~
         (P.typ_mean + condition + scaledtrialnumber) + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

model_f2c <-
  MSA_finale %>%
  lmer(., formula=f2_time_0_5 ~ 
         (P.typ_mean + condition + scaledtrialnumber) + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

model_ampc <-
  MSA_finale %>%
  lmer(., formula=maxIntensity ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute + controlIntensity +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

model_pitchc <-
  MSA_finale %>%
    mutate(meanPitch = as.integer(meanPitch)) %>%
  lmer(., formula=meanPitch ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute + controlPitch +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

model_durSegc <-
  MSA_finale %>%
  lmer(., formula=Target.segment.duration ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

# schwa offglide analysis
shmodel_f1c <-
MSA_finalschwa %>%
  lmer(., formula=f1_time_0_5 ~
         (P.typ_mean + condition + scaledtrialnumber) + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

shmodel_f2c <-
  MSA_finalschwa %>%
  lmer(., formula=f2_time_0_5 ~ 
         (P.typ_mean + condition + scaledtrialnumber) + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

shmodel_ampc <-
  MSA_finalschwa %>%
  lmer(., formula=maxIntensity ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute + controlIntensity +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

shmodel_pitchc <-
  MSA_finalschwa %>%
    mutate(meanPitch = as.integer(meanPitch)) %>%
  lmer(., formula=meanPitch ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute + controlPitch +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

shmodel_durSegc <-
  MSA_finalschwa %>%
  lmer(., formula=Target.segment.duration ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

# schwa + schwa offglide analysis
combo_model_DiphDurc <-
  MSA_both %>%
  lmer(., formula=DiphDur ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

combo_model_EDc <-
  MSA_both %>%
  lmer(., formula=ED ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

combo_model_DurRatc <-
  MSA_both %>%
  lmer(., formula=DurRat ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

# model word duration and pause duration
model_durWordc <-
  MSA_finale %>%
  lmer(., formula=Target.word.duration ~ 
         (P.typ_mean + condition + scaledtrialnumber)   + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

model_Pausec <-
  MSA_finale %>%
  lmer(., formula=Pause.duration ~ 
         (P.typ_mean + condition + scaledtrialnumber)  + syllablesPerMinute +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

# speech rate (focuses on an interaction effect)
combo_model_rated <-
  MSA_both %>%
  lmer(., formula=syllablesPerMinute ~ 
         condition * P.typ_mean + scaledtrialnumber +
         (1 | Transcript) + 
         (1 | P.Adj),
       verbose = 0)

```

```{r alosa_atun_refitting_TR}

# standardize datasets for schwa analyses
MSA_finale <- MSA_finale %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         Target.word.duration_s = (Target.word.duration - mean(Target.word.duration, na.rm = TRUE)) / sd(Target.word.duration,  na.rm = TRUE),
         Pause.duration_s = (Pause.duration - mean(Pause.duration, na.rm = TRUE)) / sd(Pause.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

# standardize datasets for schwa offglide analyses
MSA_finalschwa <- MSA_finalschwa %>% 
  mutate(
         # scale continuous variables
         f1_time_0_5_s = (f1_time_0_5 - mean(f1_time_0_5, na.rm = TRUE)) / sd(f1_time_0_5,  na.rm = TRUE),
         f2_time_0_5_s = (f2_time_0_5 - mean(f2_time_0_5, na.rm = TRUE)) / sd(f2_time_0_5,  na.rm = TRUE),
         maxIntensity_s = (maxIntensity - mean(maxIntensity, na.rm = TRUE)) / sd(maxIntensity,  na.rm = TRUE),
         meanPitch = as.integer(meanPitch),
         meanPitch_s = (meanPitch - mean(meanPitch, na.rm = TRUE)) / sd(meanPitch,  na.rm = TRUE),
         Target.segment.duration_s = (Target.segment.duration - mean(Target.segment.duration, na.rm = TRUE)) / sd(Target.segment.duration,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         controlPitch_s = (controlPitch - mean(controlPitch, na.rm = TRUE)) / sd(controlPitch,  na.rm = TRUE),
         controlIntensity_s = (controlIntensity - mean(controlIntensity, na.rm = TRUE)) / sd(controlIntensity,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

# standardize datasets for schwa + schwa offglide analyses
MSA_both <- MSA_both %>% 
  mutate(
         # scale continuous variables
         DiphDur_s = (DiphDur - mean(DiphDur, na.rm = TRUE)) / sd(DiphDur,  na.rm = TRUE),
         ED_s = (ED - mean(ED, na.rm = TRUE)) / sd(ED,  na.rm = TRUE),
         DurRat_s = (DurRat - mean(DurRat, na.rm = TRUE)) / sd(DurRat,  na.rm = TRUE),
         syllablesPerMinute_s = (syllablesPerMinute - mean(syllablesPerMinute, na.rm = TRUE)) / sd(syllablesPerMinute,  na.rm = TRUE),
         effect_con = (P.typ_mean - mean(P.typ_mean, na.rm = TRUE)) / sd(P.typ_mean,  na.rm = TRUE),
         # sum-code categorical predictors
         condition_s = as.factor(condition),
         condition_s = C(condition_s, sum)
  )

# run brms equivalent for schwa
alosa_atun_4_f1schwa_con <- brm(
  f1_time_0_5_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  data = MSA_finale, 
  seed = 111,
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_4_f1schwa_con"
  )

alosa_atun_5_f2schwa_con <- brm(
  f2_time_0_5_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_5_f2schwa_con"
  )

alosa_atun_2_intschwa_con <- brm(
  maxIntensity_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s + controlIntensity_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_2_intschwa_con"
  )

alosa_atun_1_f0schwa_con <- brm(
  meanPitch_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s + controlPitch_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_1_f0schwa_con"
  )

alosa_atun_3_durschwa_con <- brm(
  Target.segment.duration_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_3_durschwa_con"
  )

# run brms equivalent for schwa offglide
alosa_atun_9_f1off_con <- brm(
  f1_time_0_5_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finalschwa, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_9_f1off_con"
  )

alosa_atun_10_f2off_con <- brm(
  f2_time_0_5_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finalschwa, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_10_f2off_con"
  )

alosa_atun_7_intoff_con <- brm(
  maxIntensity_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s + controlIntensity_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finalschwa, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_7_intoff_con"
  )

alosa_atun_6_f0off_con <- brm(
  meanPitch_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s + controlPitch_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finalschwa, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_6_f0off_con"
  )

alosa_atun_8_duroff_con <- brm(
  Target.segment.duration_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finalschwa, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_8_duroff_con"
  )

# run brms equivalent for schwa + schwa offglide analysis
alosa_atun_11_durboth_con <- brm(
  DiphDur_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_both, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_11_durboth_con"
  )

alosa_atun_12_durrelboth_con <- brm(
  DurRat_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_both, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_12_durrelboth_con"
  )

alosa_atun_13_EDboth_con <- brm(
  ED_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_both, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_13_EDboth_con"
  )

# run brms equivalent for word duration
alosa_atun_14_worddur_con <- brm(
  Target.word.duration_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_14_worddur_con"
  )

# run brms equivalent for pause duration
alosa_atun_15_pausedur_con <- brm(
  Pause.duration_s ~ 
         (effect_con + condition_s + scaledtrialnumber) + syllablesPerMinute_s +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_finale, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_15_pausedur_con"
  )

# run brms equivalent for speech rate
alosa_atun_16_speechrate_con <- brm(
  syllablesPerMinute_s ~ 
         effect_con * condition_s + scaledtrialnumber +
         (1 | Transcript) + 
         (1 | P.Adj), 
  seed = 111,
  data = MSA_both, 
  cores = 4,  
  file = "./data/analyses/models/alosa_atun_16_speechrate_con"
  )

```

# pervagor_meeki

```{r pervagor_meeki_setup}

data = read.csv(here("data", "analyses", "tr", "pervagor_meeki", "Data", "ManySpeech_Praat_merged_data.csv"), header=T)

data$Duration_ms = data$Duration*1000
data_to_analyze <- data

########## SET CONTRASTS ############

###### set sum contrasts (compare to grand mean)
data_to_analyze$typicality = factor(data_to_analyze$typicality)

contrasts(data_to_analyze$typicality) = contr.treatment(3)

########### REMOVE OUTLIERS ###########

# TRIMMING >3 SD AND <3 SD, by SPEAKER

########### MEAN PITCH (semitones) ##################
data_meanf0_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Mean_f0_st), stdev = sd(Mean_f0_st)) %>%  
  filter(Mean_f0_st <= (3*stdev)+avg & 
           Mean_f0_st >= avg-(3*stdev)) %>%  
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### MAX PITCH (semitones) ##################
data_maxf0_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Max_f0_st), stdev = sd(Max_f0_st)) %>%  
  filter(Max_f0_st <= (3*stdev)+avg & 
           Max_f0_st >= avg-(3*stdev)) %>%  
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### INTENSITY (Praat measurements) ##################
data_int_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Overall_Intensity), stdev = sd(Overall_Intensity)) %>%  
  filter(Overall_Intensity <= (3*stdev)+avg & 
           Overall_Intensity >= avg-(3*stdev)) %>% 
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

########### DURATION ##################
data_dur_trim <- data_to_analyze %>% group_by(speaker) %>%
  mutate(avg = mean(Duration_ms), stdev = sd(Duration_ms)) %>%  
  filter(Duration_ms <= (3*stdev)+avg & #slow RTs greater than 3 SD
           Duration_ms >= avg-(3*stdev)) %>%  #fast RTs less than 3 SD
  select(colnames(data_to_analyze)) %>%
  as.data.frame()

##################### MODELS ############################

#set which contrast is dropped from the models

#run models twice:
#base = 3: drop typical
#base = 2: drop medium
base_n = 3

########### MEAN PITCH (semitones) ##################

contrasts(data_meanf0_trim$typicality) = contr.treatment(3,base=base_n)
contrasts(data_meanf0_trim$typicality)

```

```{r pervagor_meeki_models}

# mean F0
  #NOUN MODEL
  summary( meanf0_mod <- lmer (Mean_f0_st ~ typicality
                               + (typicality | speaker),
                               data = data_meanf0_trim[data_meanf0_trim$word_type == "target_name",],
                               control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( meanf0_mod <- lmer (Mean_f0_st ~ typicality
                               + (1 | speaker)
                               + (1 | target_word),
                               data = data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",],
                               control=lmerControl(optimizer="bobyqa")))

# max F0
   #NOUN MODEL
   summary( maxf0_mod <- lmer (Max_f0_st ~ typicality
                              + (1 | speaker),
                              data = data_maxf0_trim[data_maxf0_trim$word_type == "target_name",],
                              control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( maxf0_mod <- lmer (Max_f0_st ~ typicality
                              + (1 | speaker)
                              + (1 | target_word),
                              data = data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",],
                              control=lmerControl(optimizer="bobyqa")))

# int
  #NOUN MODEL
  summary( int_mod <- lmer (Overall_Intensity ~ typicality
                            + (1 | speaker),
                            data = data_int_trim[data_int_trim$word_type == "target_name",],
                            control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( int_mod <- lmer (Overall_Intensity ~ typicality
                            + (1 | speaker)
                            + (typicality | target_word),
                            data = data_int_trim[data_int_trim$word_type == "target_colour",],
                            control=lmerControl(optimizer="bobyqa")))

# duration  
  #NOUN MODEL
  summary( dur_mod <- lmer (Duration_ms ~ typicality
                            + (1 | speaker),
                            data = data_dur_trim[data_dur_trim$word_type == "target_name",],
                            control=lmerControl(optimizer="bobyqa")))

  #ADJECTIVE MODEL
  summary( dur_mod <- lmer (Duration_ms ~ typicality
                            + (1 | speaker)
                            + (typicality | target_word),
                            data = data_dur_trim[data_dur_trim$word_type == "target_colour",],
                            control=lmerControl(optimizer="bobyqa")))
```

```{r pervagor_meeki_refitting}

# scale
data_int_trim <- data_int_trim %>% 
  mutate(
    Overall_Intensity_s = (Overall_Intensity - mean(Overall_Intensity, na.rm = TRUE)) / sd(Overall_Intensity,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_meanf0_trim <- data_meanf0_trim %>% 
  mutate(
    Mean_f0_st_s = (Mean_f0_st - mean(Mean_f0_st, na.rm = TRUE)) / sd(Mean_f0_st,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_maxf0_trim <- data_maxf0_trim %>% 
  mutate(
    Max_f0_st_s = (Max_f0_st - mean(Max_f0_st, na.rm = TRUE)) / sd(Max_f0_st,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

data_dur_trim <- data_dur_trim %>% 
  mutate(
    Duration_ms_s = (Duration_ms - mean(Duration_ms, na.rm = TRUE)) / sd(Duration_ms,  na.rm = TRUE),
    effect_cat = as.factor(typicality),
    effect_cat = C(effect_cat, treatment)
  )

pervagor_meeki_1_intnoun_cat <- brm(
  Overall_Intensity_s ~ 
         effect_cat + 
         (1|speaker),
  seed = 111,
  data = data_int_trim[data_int_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_1_intnoun_cat"
  )

pervagor_meeki_2_durnoun_cat <- brm(
  Duration_ms_s ~ 
         effect_cat + 
         + (1| speaker),
  seed = 111,
  data = data_dur_trim[data_dur_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_2_durnoun_cat"
  )

pervagor_meeki_3_f0meannoun_cat <- brm(
  Mean_f0_st_s ~ 
         effect_cat + 
         + (effect_cat | speaker),
  seed = 111,
  data = data_meanf0_trim[data_meanf0_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_3_f0meannoun_cat"
  )

pervagor_meeki_4_f0maxnoun_cat <- brm(
  Max_f0_st_s ~ 
         effect_cat + 
         + (1| speaker),
  seed = 111,
  data = data_maxf0_trim[data_maxf0_trim$word_type == "target_name",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_4_f0maxnoun_cat"
  )

pervagor_meeki_5_intadj_cat <- brm(
  Overall_Intensity_s ~ 
         effect_cat 
         + (1 | speaker)
         + (effect_cat | target_word),
  seed = 111,
  data = data_int_trim[data_int_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_5_intadj_cat"
  )

pervagor_meeki_6_duradj_cat <- brm(
  Duration_ms_s ~ 
         effect_cat + 
          + (1 | speaker)
          + (effect_cat | target_word),
  seed = 111,
  data = data_dur_trim[data_dur_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_6_duradj_cat"
  )

pervagor_meeki_7_f0meanadj_cat <- brm(
  Mean_f0_st_s ~ 
         effect_cat + 
         + (1 | speaker)
         + (1 | target_word),
  seed = 111,
  data = data_meanf0_trim[data_meanf0_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_7_f0meanadj_cat"
  )

pervagor_meeki_8_f0maxadj_cat <- brm(
  Max_f0_st_s ~ 
         effect_cat + 
         + (1 | speaker)
         + (1 | target_word),
  seed = 111,
  data = data_maxf0_trim[data_maxf0_trim$word_type == "target_colour",], 
  cores = 4,  
  file = "./data/analyses/models/pervagor_meeki_8_f0maxadj_cat"
  )

```

# procambarus_maculosus

```{r procambarus_maculosus_setup}

getPCscores <- function(fpcaObj) {
  if (is.matrix(fpcaObj$scores)) {
    fpcaObj$scores
  } else if (is.array(fpcaObj$scores) & length(dim(fpcaObj$scores)) == 3) {
    apply(fpcaObj$scores, 1:2, sum)
  } else {
    stop("getPCscores: invalid input fpcaObj")
  }
}

##### load data #####

data <- read.delim("./data/analyses/tr/procambarus_maculosus/selectedData.txt", header = T, sep = "") %>% 
  as_tibble() # %>% 
  # filter(sex == "female") # only run on female data!

# data = read.csv("./data/analyses/tr/pervagor_meeki/Data/ManySpeech_Praat_merged_data.csv", header=T)

info <- data %>% 
  dplyr::select(sl_rowIdx, bundle, vowel, speaker, sex, typicality, word) %>% 
  unique()

##### FPCA #####

# set parameters
nKnots <- 6
lambda <- 1e-8
Lfdobj <- 2
nOrder <- 2 + Lfdobj
nBasis <- nKnots + nOrder - 2

# create a basis for B-splines
basis <- fda::create.bspline.basis(rangeval = c(0, 1), nbasis = nBasis, norder = nOrder)

# define functional parameter object (will be used to estimate functional parameters later)
fdParObj <- fda::fdPar(fdobj = basis, Lfdobj = Lfdobj, lambda = lambda)

coef <- data %>%
  group_by(sl_rowIdx) %>%
  summarise(coef = c(fda::smooth.basis(argvals = times_norm, y = F0, fdParObj)$fd$coefs),
            coefId = seq_len(nBasis)) %>%
  ungroup() %>%
  pivot_wider(names_from = coefId, values_from = coef) %>%
  select(-sl_rowIdx) %>% 
  abind::abind(along = 2) %>%
  aperm(c(2,1))

# check class and dimension
coef %>% class() # array
coef %>% dim()   # 8 842; 8 671

# create functional data object
y_fd <- fda::fd(coef = coef, basisobj = basis)

# this is the actual FPCA
nPCs <- 3
y_pcafd <- fda::pca.fd(y_fd, nharm = nPCs, fdParObj)
# saveRDS(y_pcafd, file = "./pcafd.rds")

y_pcafd$varprop 
# both sexes: 0.92432242 0.04662226 0.01509360
# only female: 0.87031521 0.08118398 0.02410822

info %<>% 
  cbind(y_pcafd %>% getPCscores) %>%
  rename(s1 = `1`, s2 = `2`, s3 = `3`) %>% 
  as_tibble()
# write.table(info, "./PCscores.txt")

##### interpretation #####

tx <- seq(1/1000, 1, by = 1/1000)

curves <- expand.grid(time = tx,
                      PC = 1:3,
                      perturbation = c(-1, 0, 1)
) %>% data.table::setDT()

sd_score <- info %>% summarise_at(vars(s1:s3), sd) %>% as.numeric()

curves[, value := (y_pcafd$meanfd$coefs[,1] + perturbation * sd_score[PC] * y_pcafd$harmonics$coefs[,PC]) %>% 
         fda::fd(., basis) %>%
         fda::eval.fd(tx, .),
       by = .(PC, perturbation)]

reconstruct <- curves %>% mutate(per = ifelse(perturbation == 0, "mu[F0](t)",
                                              ifelse(perturbation == 1, "mu[F0](t) + sigma[s[k]] %.% PCk[F0](t)",
                                                     "mu[F0](t) - sigma[s[k]] %.% PCk[F0](t)")),
                                 k = case_when(PC == 1 ~ "k == 1",
                                               PC == 2 ~ "k == 2",
                                               PC == 3 ~ "k == 3"))

reconstruct$per <- factor(reconstruct$per)

levels(reconstruct$per) = list("mu[F0](t) - sigma[s[k]] %.% PCk[F0](t)" = "0",
                               "mu[F0](t)" = "1",
                               "mu[F0](t) + sigma[s[k]] %.% PCk[F0](t)" = "2")

# pl <- ggplot(reconstruct) + 
#   aes(x = time, y = value) + 
#   geom_line(size = 1.2) + 
#   facet_grid(k~per, labeller = label_parsed) +
#   xlab("Normalised Time") + ylab("F0 [Hz]") + theme + 
#   guides(linetype = guide_legend(override.aes = list(size = 1)))
# ggsave(filename = "./PCcurves.png", plot = pl, width = 25, height = 20, unit = "cm")
# 
# pl <- ggplot(info %>% pivot_longer(cols = s1:s3) %>% 
#                mutate(score = case_when(name == "s1" ~ "s[1]",
#                                         name == "s2" ~ "s[2]",
#                                         name == "s3" ~ "s[3]"))) + 
#   aes(x = 0, y = value, fill = typicality) + 
#   facet_grid(score~., scales = "free_y", labeller = label_parsed, switch = "y") + 
#   geom_boxplot() + ylab("") + xlab("") +
#   theme + theme(strip.background.y = element_rect(fill = "white"), 
#                 strip.placement = "outside",
#                 legend.margin = margin(t = -15))
# ggsave(filename = "./PCscores.png", plot = pl, width = 18, height = 18, unit = "cm")

##### different interpretation #####

tx <- seq(1/1000, 1, by = 1/1000)

curves <- expand.grid(time = tx,
                      PC = 1:2,
                      perturbation = seq(-1, 1, by = .25)
) %>% data.table::setDT()

sd_score <- info %>% summarise_at(vars(s1:s3), sd) %>% as.numeric()

curves[, value := (y_pcafd$meanfd$coefs[,1] + perturbation * sd_score[PC] * y_pcafd$harmonics$coefs[,PC]) %>% 
         fda::fd(., basis) %>%
         fda::eval.fd(tx, .),
       by = .(PC, perturbation)]

PC_labeller <- as_labeller(function(x) paste0('PC', x))

# pl <- ggplot(curves) +
#   aes(x = time, y = value, col = perturbation, group = perturbation) +
#   geom_line(size = 1.2) +
#   scale_color_gradient2(name = "P ", low = "cornflowerblue", mid = "grey", high = "darkgoldenrod1") +
#   facet_grid(~PC, labeller = PC_labeller) +
#   geom_line(data = curves[perturbation == 0], color = "black", size = 1.5) +
#   xlab("Normalised Time") + ylab("F0 [Hz]") +
#   theme(legend.direction = "horizontal", 
#                 legend.title = element_text(),
#                 legend.text = element_text(size = 12))
# ggsave(filename = "./PCcurves_col.png", plot = pl, width = 20, height = 12, unit = "cm")

```

```{r procambarus_maculosus_models}

s1.lmer <- lmer(s1 ~ typicality * sex + (sex | word) + (typicality | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) %>% 
  step() %>% get_model()

# they did not interpret the models with typicality in there, so the minimum model we refit will be typicality + sex

s2.lmer <- lmer(s2 ~ typicality * sex + (sex | word) + (typicality | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) %>% 
  step() %>% get_model()
# s2 ~ typicality + (typicality | speaker), but failed to converge

s2.lmer <- lmer(s2 ~ typicality + (1 | word) + (1 | speaker), data = info,
                control = lmerControl(optimizer = "optimx", optCtrl = list(method = "nlminb"))) 
s2.lmer %>% summary(corr = F)

```

```{r procambarus_maculosus_refitting}

dat <- info %>% 
  mutate(
         # scale continuous variables
         s1_s = (s1 - mean(s1, na.rm = TRUE)) / sd(s1,  na.rm = TRUE),
         s2_s = (s2 - mean(s2, na.rm = TRUE)) / sd(s2,  na.rm = TRUE),
         sex_s = as.factor(sex),
         sex_s = C(sex_s, sum),
         effect_cat = as.factor(typicality),
         effect_cat = C(effect_cat, treatment)
  )

# run brms on standardized models
procambarus_maculosus_1_pca1_cat <- brm(
  s1_s ~ 
         effect_cat + sex_s + 
    (sex_s | word) + 
    (effect_cat | speaker),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_maculosus_1_pca1_cat"
  )

# run brms on standardized models
procambarus_maculosus_2_pca2_cat <- brm(
  s2_s ~ 
         effect_cat + 
    (1 | word) +
    (1 | speaker),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_maculosus_2_pca2_cat"
  )

```

# dunkleosteus_inscriptus

```{r dunkleosteus_inscriptus_setup}

path <- here("data", "analyses", "tr", "dunkleosteus_inscriptus", "data/")

flist_times  <- list.files(path = here("data", "analyses", "tr", "dunkleosteus_inscriptus", "data"), pattern = "^timings(.*)xlsx")
get_times    <- lapply(paste0(path, flist_times), read_excel)
times        <- bind_rows(get_times)

flist_trials <- list.files(path = here("data", "analyses", "tr", "dunkleosteus_inscriptus", "data"), pattern = "^trials(.*)xlsx")
get_trials   <- lapply(paste0(path, flist_trials), read_excel)
trials       <- bind_rows(get_trials)

mfields      <- c( "trial","speaker" )
dat          <- merge( trials, times, by=mfields )

dat$condition      <- factor(dat$condition)
dat$typicality     <- factor(dat$typicality)
dat$target_name    <- factor(dat$target_name)
dat$target_colour  <- factor(dat$target_colour)

# correct constituent names exported by <extract_timings_MSA.m> and 
# calculate additional durations derived from from Praat intervals
dat$dur_n    <- dat$dur_np
dat$dur_np   <- dat$dur_ap
dat$dur_a    <- dat$dur_np - dat$dur_n
dat$dur_ap   <- NULL
dat$rel_a    <- dat$dur_a / dat$dur_dp
dat$rel_n    <- dat$dur_n / dat$dur_dp

## check extremely long utterances
dur_ip_max     <- mean(dat$dur_ip) + 2.5*sd(dat$dur_ip)
#(dat_long_ip   <- dat[ dat$dur_ip>dur_ip_max, ])
# 28 utterances longer than 3.91s rechecked with <check_maus.m>
# all 28 invalid: misproduced, contain hesitations, or are incorrectly segmented
# remove all trials longer than 2.5sd mean duration:
dat  <- dat[ !(dat$dur_ip > dur_ip_max ), ]

## check duration extrema: preposition phrase 'auf ...'
dur_pp_min     <- mean(dat$dur_pp) - 2*sd(dat$dur_pp)
#(dat_short_pp  <- dat[ dat$dur_pp < dur_pp_min, ])
# 4 trials with PPs shorter than 728ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_pp < dur_pp_min ), ]

dur_pp_max     <- mean(dat$dur_pp) + 3*sd(dat$dur_pp)
#(dat_long_pp   <- dat[ dat$dur_pp > dur_pp_max, ])
# 19 trials with PPs longer than 2.03s rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_pp > dur_pp_max ), ]

## check duration extrema: target determiner phrase
dur_dp_min     <- mean(dat$dur_dp) - 2*sd(dat$dur_dp)
#(dat_short_dp  <- dat[ dat$dur_dp < dur_dp_min, ])
# 5 trials with target DPs shorter than 629ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_dp < dur_dp_min ), ]

dur_dp_max     <- mean(dat$dur_dp) + 3.1*sd(dat$dur_dp)
#(dat_long_dp   <- dat[ dat$dur_dp > dur_dp_max, ])
# 5 trials with target DPs longer than 1.75s rechecked with <check_maus.m>
# mainly just long words/slow speech rate: don't remove from dataset

## check duration extrema: target noun phrase
dur_np_min     <- mean(dat$dur_np) - 2*sd(dat$dur_np)
#(dat_short_np  <- dat[ dat$dur_np < dur_np_min, ])
# 26 trials with target NPs shorter than 460ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_np < dur_np_min ), ]

dur_np_max     <- mean(dat$dur_np) + 3.1*sd(dat$dur_np)
#(dat_long_np   <- dat[ dat$dur_np > dur_np_max, ])
# 3 trials with target APs longer than 1.55s rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_np > dur_np_max ), ]

## check duration extrema: target nouns
dur_n_min     <- mean(dat$dur_n) - 1.8*sd(dat$dur_n)
#(dat_short_n  <- dat[ dat$dur_n < dur_n_min, ])
# 28 trials with target nouns shorter than 217ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_n < dur_n_min ), ]

dur_n_max     <- mean(dat$dur_n) + 2.5*sd(dat$dur_n)
#(dat_long_n   <- dat[ dat$dur_n > dur_n_max, ])
# 23 trials with target nouns longer than 875ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_n > dur_n_max ), ]

## check duration extrema: target adjectives
dur_a_min     <- mean(dat$dur_a) - 1.6*sd(dat$dur_a)
#(dat_short_a  <- dat[ dat$dur_a < dur_a_min, ])
# 25 trials with target adjectives shorter than 173ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_a < dur_a_min ), ]

dur_a_max     <- mean(dat$dur_a) + 3.85*sd(dat$dur_a)
#(dat_long_a   <- dat[ dat$dur_a > dur_a_max, ])
# 4 trials with target adjectives longer than 947ms rechecked with <check_maus.m>
# all involve some segmentation inaccuracies: remove from dataset
dat  <- dat[ !(dat$dur_a > dur_a_max ), ]

## ===== select trials eliciting target words in NF condition  =====
dat_nf  <- dat[ (dat$condition=='NF'), ]
dat_nf$condition   <- factor(dat_nf$condition)
dat_nf$typicality  <- factor(dat_nf$typicality)

```

```{r dunkleosteus_inscriptus_models}

## Bayesian model:
fm2 <- rel_n ~ typicality + (1|speaker) + (1|target_name) + (1|target_colour)
fm2a <- dur_n ~ typicality + (1|speaker) + (1|target_name) + (1|target_colour)
fm3 <- rel_a ~ typicality + (1|speaker) + (1|target_name) + (1|target_colour)
fm3a <- dur_a ~ typicality + (1|speaker) + (1|target_name) + (1|target_colour)
fm4  <- dur_np ~ typicality + (1|speaker) + (1|target_name) + (1|target_colour)

# TR: claimed to run fm2a and fm3a as Bayesian, but scripts do not provide these models, so I take the frequentist ones
lm2a <- lmer(fm2a, data=dat)
lm3a <- lmer(fm3a, data=dat)

# TR: They specified priors but ended up running default priors (strange), also unclear which data are used, subsetted ones ("dat_nf") or full data ("dat"). I will pick dat for all. 
pri2 <- get_prior( formula=fm2, data=dat )
pri3 <- get_prior( formula=fm3, data=dat )
pri4 <- get_prior( formula=fm4, data=dat )

# TR: add cores to fit quicker
bm2   <- brm( fm2, dat, prior=pri2, cores = 4 )
bm3   <- brm( fm3, dat, prior=pri3, cores = 4 )
bm4   <- brm( fm4, dat, prior=pri4, cores = 4 )

```

```{r dunkleosteus_inscriptus_refitting}

# standardizing 
dat <- dat %>% 
  mutate(
         # scale continuous variables
         rel_n_s = (rel_n - mean(rel_n, na.rm = TRUE)) / sd(rel_n,  na.rm = TRUE),
         rel_a_s = (rel_a - mean(rel_a, na.rm = TRUE)) / sd(rel_a,  na.rm = TRUE),
         dur_n_s = (dur_n - mean(dur_n, na.rm = TRUE)) / sd(dur_n,  na.rm = TRUE),
         dur_a_s = (dur_a - mean(dur_a, na.rm = TRUE)) / sd(dur_a,  na.rm = TRUE),
         dur_np_s = (dur_np - mean(dur_np, na.rm = TRUE)) / sd(dur_np,  na.rm = TRUE),
         effect_cat = C(typicality, treatment)
  )

# run brms on standardized models
dunkleosteus_inscriptus_1_relndur_cat <- brm(
  rel_n_s ~ 
         effect_cat + 
         (1|speaker) + 
         (1|target_name) + 
         (1|target_colour),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/dunkleosteus_inscriptus_1_relndur_cat"
  )

# did initially not converge, adjusted adapt_delta to 0.9999  
dunkleosteus_inscriptus_2_ndur_cat <- brm(
  dur_n_s ~ 
         effect_cat + 
         (1|speaker) + 
         (1|target_name) + 
         (1|target_colour),
  seed = 111,
  data = dat, 
  control = list(adapt_delta = 0.9999),
  cores = 4,  
  file = "./data/analyses/models/dunkleosteus_inscriptus_2_ndur_cat"
  )

dunkleosteus_inscriptus_3_reladur_cat <- brm(
  rel_a_s ~ 
         effect_cat + 
         (1|speaker) + 
         (1|target_name) + 
         (1|target_colour),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/dunkleosteus_inscriptus_3_reladur_cat"
  )

dunkleosteus_inscriptus_4_adur_cat <- brm(
  dur_a_s ~ 
         effect_cat + 
         (1|speaker) + 
         (1|target_name) + 
         (1|target_colour),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/dunkleosteus_inscriptus_4_adur_cat"
  )

dunkleosteus_inscriptus_5_npdur_cat <- brm(
  dur_np_s ~ 
         effect_cat + 
         (1|speaker) + 
         (1|target_name) + 
         (1|target_colour),
  seed = 111,
  data = dat, 
  cores = 4,  
  file = "./data/analyses/models/dunkleosteus_inscriptus_5_npdur_cat"
  )

```

# chelonia_brummeri

```{r chelonia_brummeri_setup}

# function to convert hz to semitone
f2st <- function(hz, base) {
  f0_st <- 12 * log2(hz / base)
  return(f0_st)
}

#################
### Load data ###
#################
d <- read_delim(here("data", "analyses", "tr", "chelonia_brummeri", "analysis", "german_msa_prosody.txt"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)
syll <- read_csv(here("data", "analyses", "tr", "chelonia_brummeri", "analysis", "syllables.csv"))

#######################
### Data processing ###
#######################
# Remove errors and hesitations
d_orig <- d
d <- subset(d, is.na(notes))

# Retain only NF condition which is the only one with typicality ratings
d <- subset(d, cond == "NF")

# label half
# d$half <- ifelse(d$nTrial <= 35, "first", "second")
# TR: DID not reproduce, needed to change nTrial to trial
d$half <- ifelse(d$trial <= 35, "first", "second")

# Get durations in ms
d$adj_vdur <- (d$adj_v_end - d$adj_v_start) * 1000
d$utt_dur <- d$utt_dur*1000

# Get relative duration (NF cond only)
d <- merge(d, syll, by = "noun")
d$avg_syll_dur <- d$utt_dur / d$n_syll
d$rel_dur <- d$adj_vdur - d$avg_syll_dur

# Convert pascal to dB
d$adj_v_intens <- 20*log10(d$adj_v_amp/0.00002)
d$utt_intens <- 20*log10(d$utt_amp/0.00002)

# Get relative intensity
d$rel_intens <- d$adj_v_intens - d$utt_intens

# Convert hertz to semitone
d$adj_maxf0 <- as.numeric(as.character(d$adj_maxf0))
d$adj_time_maxf0 <- as.numeric(as.character(d$adj_time_maxf0))
d$adj_minf0 <- as.numeric(as.character(d$adj_minf0))
d$adj_time_minf0 <- as.numeric(as.character(d$adj_time_minf0))

# get f0 min for each participant from the two min columns
bases <- c()
my_subjects <- unique(d$subj)
for (i in 1:length(my_subjects)) {
  tmp <- subset(d, subj == my_subjects[i])
  minf0 <- min(tmp$adj_minf0, na.rm = T)
  bases <- append(bases, minf0)
}

bases <- cbind(my_subjects, bases)
bases <- data.frame(bases)
colnames(bases) <- c("subj", "minf0")
d <- merge(d, bases, by = "subj")
d$minf0 <- as.numeric(as.character(d$minf0))

d$adj_maxf0_st <- f2st(d$adj_maxf0, d$minf0)
d$adj_minf0_st <- f2st(d$adj_minf0, d$minf0)
d$f0range_st <- d$adj_maxf0_st - d$adj_minf0_st

```

```{r chelonia_brummeri_models}

# treatment coding with baseline = typical
d$nTypicality1 <- ifelse(d$typicality == "atypical", 1, 0)
d$nTypicality2 <- ifelse(d$typicality == "medium", 1, 0)

# syllable structure
d$syll_structure <- ifelse(d$adj %in% c("roten"), "CV", ifelse(d$adj %in% c("gruenen", "braunen"), "CCV", "CVC"))
d$nSyll1 <- ifelse(d$syll_structure == "CCV", 1, ifelse(d$syll_structure == "CV", -1, 0))
d$nSyll2 <- ifelse(d$syll_structure == "CVC", 1, ifelse(d$syll_structure == "CV", -1, 0))

# center DVs
d$Nreldur <- d$rel_dur - mean(d$rel_dur)
d$Nrelintens <- d$rel_intens - mean(d$rel_intens)
d$Nadj_maxf0_st <- d$adj_maxf0_st - mean(d$adj_maxf0_st)
d$Nf0range_st <- d$f0range_st - mean(d$f0range_st)
d$Nadj_maxf0 <- d$adj_maxf0 - mean(d$adj_maxf0)
# TR: does not reproduce
# d$Nadj_maxf0_erb <- d$adj_maxf0_erb - mean(d$adj_maxf0_erb)
# TR: I assume, the authors meant st?
d$Nf0range_st <- d$f0range_st - mean(d$f0range_st)
d$nTrial <- d$trial - 35

# run models 
# TR: added cores to speed up
fit_reldur2 <- brm(Nreldur ~ nTypicality1*nTrial + nTypicality2*nTrial + nSyll1 + nSyll2 + (1 + nTypicality1  + nTypicality2 |subj) + (1 | adj), d, control = list(adapt_delta = 0.99), cores = 4, prior = c(
  prior(normal(0, 200), class = Intercept),
  prior(normal(0, 200), class = b, coef = nTypicality1),
  prior(normal(0, 200), class = b, coef = nTypicality2),
  prior(normal(0, 200), class = b, coef = nTrial), 
  prior(normal(0, 200), class = b, coef = nSyll1),
  prior(normal(0, 200), class = b, coef = nSyll2))
)

fit_relintens2 <- brm(Nrelintens ~ nTypicality1*nTrial + nTypicality2*nTrial + nSyll1 + nSyll2 + (1 + nTypicality1  + nTypicality2 |subj) + (1 | adj), d, control = list(adapt_delta = 0.99), cores = 4, prior = c(
  prior(normal(0, 4), class = Intercept), 
  prior(normal(0, 4), class = b, coef = nTypicality1),
  prior(normal(0, 4), class = b, coef = nTypicality2), 
  prior(normal(0, 4), class = b, coef = nTrial),
  prior(normal(0, 4), class = b, coef = nSyll1),
  prior(normal(0, 4), class = b, coef = nSyll2))
)

fit_maxf0st2 <- brm(Nadj_maxf0_st ~ nTypicality1*nTrial + nTypicality2*nTrial + nSyll1 + nSyll2 + (1 + nTypicality1  + nTypicality2 |subj) + (1 | adj), d, control = list(adapt_delta = 0.9999, cores = 4, max_treedepth = 15), iter = 4000, prior = c(
  prior(normal(0, 7), class = Intercept),
  prior(normal(0, 7), class = b, coef = nTypicality1),
  prior(normal(0, 7), class = b, coef = nTypicality2),
  prior(normal(0, 7), class = b, coef = nTrial),
  prior(normal(0, 7), class = b, coef = nSyll1),
  prior(normal(0, 7), class = b, coef = nSyll2))
)

fit_f0range2 <- brm(Nf0range_st ~ nTypicality1*nTrial + nTypicality2*nTrial + nSyll1 + nSyll2 + (1 + nTypicality1  + nTypicality2 |subj) + (1 | adj), d, control = list(adapt_delta = 0.9999, cores = 4, max_treedepth = 15), iter = 4000, prior = c(
  prior(normal(0, 7), class = Intercept),
  prior(normal(0, 7), class = b, coef = nTypicality1),
  prior(normal(0, 7), class = b, coef = nTypicality2),
  prior(normal(0, 7), class = b, coef = nTrial),
  prior(normal(0, 7), class = b, coef = nSyll1),
  prior(normal(0, 7), class = b, coef = nSyll2))
)

```

```{r chelonia_brummeri_refitting}

# standardize
d <- d %>% 
  mutate(
    Nreldur_s = (Nreldur - mean(Nreldur, na.rm = TRUE)) / sd(Nreldur,  na.rm = TRUE),
    Nrelinten_s = (Nrelintens - mean(Nrelintens, na.rm = TRUE)) / sd(Nrelintens,  na.rm = TRUE),
    Nadj_maxf0_st_s = (Nadj_maxf0_st - mean(Nadj_maxf0_st, na.rm = TRUE)) / sd(Nadj_maxf0_st,  na.rm = TRUE),
    Nf0range_st_s = (Nf0range_st - mean(Nf0range_st, na.rm = TRUE)) / sd(Nf0range_st,  na.rm = TRUE),
    nTrial_s = (nTrial - mean(nTrial, na.rm = TRUE)) / sd(nTrial,  na.rm = TRUE),
    effect_cat = C(as.factor(typicality), treatment),
    syll_structure_s = C(as.factor(syll_structure), sum)
  )

# refit models 
# TR: instead of hard coding treatment coding as the authors did, we simplify the model by using deviation coding for typicality
# duration
chelonia_brummeri_1_dur_cat <- brm(
  Nreldur_s ~ 
         effect_cat*nTrial_s + 
         syll_structure_s + 
         (1 + effect_cat |subj) + 
         (1 | adj),
  seed = 111,
  data = d, 
  cores = 4,  
  file = "./data/analyses/models/chelonia_brummeri_1_dur_cat"
  )

# intensity
chelonia_brummeri_2_int_cat <- brm(
  Nrelinten_s ~ 
           effect_cat*nTrial_s + 
         syll_structure_s + 
         (1 + effect_cat |subj) + 
         (1 | adj),
  seed = 111,
  data = d, 
  cores = 4,  
  file = "./data/analyses/models/chelonia_brummeri_2_int_cat"
  )

# f0 range
chelonia_brummeri_3_f0range_cat <- brm(
  Nf0range_st_s ~ 
         effect_cat*nTrial_s + 
         syll_structure_s + 
         (1 + effect_cat |subj) + 
         (1 | adj),
  seed = 111,
  data = d, 
  cores = 4,  
  file = "./data/analyses/models/chelonia_brummeri_3_f0range_cat"
  )

# f0 max
chelonia_brummeri_4_f0max_cat <- brm(
  Nadj_maxf0_st_s ~ 
          effect_cat*nTrial_s + 
         syll_structure_s + 
         (1 + effect_cat |subj) + 
         (1 | adj),
  seed = 111,
  data = d, 
  cores = 4,  
  file = "./data/analyses/models/chelonia_brummeri_4_f0max_cat"
  )

```

# procambarus_mahogoni

```{r procambarus_mahogoni_setup}

# read csv
msa_data_raw <- read_csv(here("data", "analyses", "tr", "procambarus_mahogoni", "MSA-data.csv"))

# transform raw data 
msa_data <- msa_data_raw %>%
mutate(v1_duration_c_std = (v1_duration_c - mean(v1_duration_c)) / sd(v1_duration_c),
       v1_pitch_mean_c_std = (v1_pitch_mean_c - mean(v1_pitch_mean_c)) / sd(v1_pitch_mean_c),
       v1_int_mean_c_std = (v1_int_mean_c - mean(v1_int_mean_c)) / sd(v1_int_mean_c))

```

```{r procambarus_mahogoni_models}

mod1 <- lmer(v1_duration_c_std ~ typ_median +
                     (1|participant) + (1|typ_median), data = msa_data)

mod2 <- lmer(v1_pitch_mean_c_std ~ typ_median +
               (1|participant) + (1|typ_median), data = msa_data)

mod3 <- lmer(v1_int_mean_c_std ~ typ_median +
               (1|participant) + (1|typ_median), data = msa_data)

```

```{r procambarus_mahogoni_refitting}

# standardize
msa_data <- msa_data %>% 
  mutate(effect_con = (typ_median - mean(typ_median, na.rm = TRUE)) / sd(typ_median,  na.rm = TRUE))

procambarus_mahogoni_1_dur_con <- brm(
  v1_duration_c_std ~ 
         effect_con + 
         (1|participant) + 
         (1|effect_con),
  seed = 111,
  data = msa_data, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_mahogoni_1_dur_con"
  )

procambarus_mahogoni_2_f0_con <- brm(
  v1_pitch_mean_c_std ~ 
         effect_con + 
         (1|participant) + 
         (1|effect_con),
  seed = 111,
  data = msa_data, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_mahogoni_2_f0_con"
  )

procambarus_mahogoni_3_int_con <- brm(
  v1_int_mean_c_std ~ 
         effect_con + 
         (1|participant) + 
         (1|effect_con),
  seed = 111,
  data = msa_data, 
  cores = 4,  
  file = "./data/analyses/models/procambarus_mahogoni_3_int_con"
  )

```

# pseudopleuronectes_assasi

```{r pseudopleuronectes_assasi_setup}

# read csv
mpa <- read.delim("./data/analyses/tr/pseudopleuronectes_assasi/alldata.txt", header = T, sep = "") %>% 
  as_tibble()

atypical <- subset(mpa,Typicality=="atypical")
atypicalAF <- subset(atypical,Condition=="AF")
atypicalNF <- subset(atypical,Condition=="NF")
atypicalAF1 <- subset(atypicalAF,Vowel==1)
atypicalAF2 <- subset(atypicalAF,Vowel==2)
atypicalNF1 <- subset(atypicalNF,Vowel==1)
atypicalNF2 <- subset(atypicalNF,Vowel==2)
atypical1 <- subset(atypical,Vowel==1)
atypical2 <- subset(atypical,Vowel==2)
atypicalAFNF1 <- rbind(atypicalAF1,atypicalNF1)
NF <- subset(mpa,Condition=="NF")
NFa <- subset(NF,Typicality=="atypical")
NFm <- subset(NF,Typicality=="medium")
NFt <- subset(NF,Typicality=="typical")
NF1 <- subset(NF,Vowel==1)
NF2 <- subset(NF,Vowel==2)
```

```{r pseudopleuronectes_assasi_models}

NF1DurTyp <- lmer(Duration~Typicality+(1|Speaker)+(1|TargetColor),data=NF1,REML=FALSE)
NF1F0Typ <- lmer(MedianST~Sex+Typicality+(1|Speaker)+(1|TargetColor),data=NF1,REML=FALSE)
NF1IntTyp <- lmer(MedianInt~Typicality+(1|Speaker)+(1|TargetColor),data=NF1,REML=FALSE)
NF1DistTyp <- lmer(FormDist~Typicality+(1|Speaker)+(1|TargetColor),data=NF1,REML=FALSE)

```

```{r pseudopleuronectes_assasi_refitting}

# standardize
NF1 <- NF1 %>% 
    mutate(
    Duration_s = (Duration - mean(Duration, na.rm = TRUE)) / sd(Duration,  na.rm = TRUE),
    MedianST_s = (MedianST - mean(MedianST, na.rm = TRUE)) / sd(MedianST,  na.rm = TRUE),
    MedianInt_s = (MedianInt - mean(MedianInt, na.rm = TRUE)) / sd(MedianInt,  na.rm = TRUE),
    FormDist_s = (FormDist - mean(FormDist, na.rm = TRUE)) / sd(FormDist,  na.rm = TRUE),
    effect_cat = C(as.factor(Typicality), treatment),
    sex_s = as.factor(Sex),
    sex_s = C(sex_s, sum)
  )

pseudopleuronectes_assasi_1_dur_cat <- brm(
  Duration_s ~ 
         effect_cat + 
         (1|Speaker) + 
         (1|TargetColor),
  seed = 111,
  data = NF1, 
  cores = 4,  
  file = "./data/analyses/models/pseudopleuronectes_assasi_1_dur_cat"
  )

pseudopleuronectes_assasi_2_f0_cat <- brm(
  MedianST ~ 
         effect_cat + Sex +
         (1|Speaker) + 
         (1|TargetColor),
  seed = 111,
  data = NF1, 
  cores = 4,  
  file = "./data/analyses/models/pseudopleuronectes_assasi_2_f0_cat"
  )

pseudopleuronectes_assasi_3_int_cat <- brm(
  FormDist_s ~ 
         effect_cat + 
         (1|Speaker) + 
         (1|TargetColor),
  seed = 111,
  data = NF1, 
  cores = 4,  
  file = "./data/analyses/models/pseudopleuronectes_assasi_3_int_cat"
  )

pseudopleuronectes_assasi_4_formants_cat <- brm(
  MedianInt_s ~ 
         effect_cat + 
         (1|Speaker) + 
         (1|TargetColor),
  seed = 111,
  data = NF1, 
  cores = 4,  
  file = "./data/analyses/models/pseudopleuronectes_assasi_4_formants_cat"
  )

```