### 2023-01-18 data/analyses/README.Rmd
update all READMEs

---
title: "Analyses"
output: github_document
date: 'Last update: `r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

The `analyses` directory contains all of the analytic teams' data (downloaded from OSF). 
The teams' analyses were refit by the initiating authors (sc, jvc, and tr) and the corresponding rds files are stored in the `models` directory. 
The file `msa_mean_sd_df.csv` contains the unstandardized mean and standard deviations of every submitted model containing pitch, duration or intensity as the outcome measure. 
Finally, `msa_models.rds` contains the coding sheet completed by the initiating authors of every submitted analysis.

```{r, 'analysis-structure', comment=NA, echo=F}
fs::dir_tree(recurse = 0)
```